### 2019-10-22 ms/ms.tex
First draft

\documentclass[12pt]{report}
\usepackage[utf8]{inputenc}
\usepackage{times}
\renewcommand\thesection{\arabic{section}}
\usepackage[parfill]{parskip}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage[left=0.75in,right=0.75in,top=0.75in,bottom=0.75in]{geometry}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{nth}
\usepackage{bm}
\usepackage{siunitx}
\usepackage[labelfont=bf]{caption}
\usepackage[doublespacing]{setspace}
\usepackage{filecontents}
\usepackage{lineno} \linenumbers
\usepackage{amsmath}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Bibliography%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[style=numeric-comp,sorting=none,uniquelist=minyear,uniquename=false,maxcitenames=2,maxbibnames=5,doi=false,url=false,isbn=false,defernumbers=true,giveninits=true]{biblatex}
%Italicise et al.
\renewbibmacro*{name:andothers}{
% Based on name:andothers from biblatex.def
\ifboolexpr{ test{\ifnumequal{\value{listcount}}{\value{liststop}}} and test \ifmorenames } {\ifnumgreater{\value{liststop}}{1} {\finalandcomma}
{}%
\andothersdelim\bibstring[\emph]{andothers}} {}}
%Make 'and' an ampersand
\renewcommand*{\finalnamedelim}{%
\ifnumgreater{\value{liststop}}{2}{\finalandcomma}{}%
\addspace\&\space}
%Remove 'in:'
\renewbibmacro{in:}{}
%Remove pp. on pages numbers
\DeclareFieldFormat[article]{pages}{#1}%
%Remove unwanted entries
\AtEveryBibitem{%
  \clearfield{day}%
  \clearfield{month}%
  \clearfield{endday}%
  \clearfield{endmonth}%
  \clearfield{url}%
  \clearfield{URL}%
  \clearfield{eprint}%
}
%Remove spaces between initials
%Remove double quotes around titles
\DeclareFieldFormat*{citetitle}{#1}
\DeclareFieldFormat*{title}{#1}
\addbibresource{/home/will/Papers/Library.bib}
\begin{document}
\section*{Title page}
\textbf{Article title}: Placing species' climate envelope within its
historic context reveals remarkable idiosyncrasy of distributional
change

\textbf{Authors:} William D.\ Pearse$^{1,*}$ \& T.\ Jonathan
Davies$^{2,*}$

$^1$ Department of Biology \& Ecology Center, Utah State University,
5305 Old Main Hill, Logan UT, 84322

$^2$ Department of Biological Sciences, Universit\'{e} du Qu\'{e}bec
\`{a} Montr\'{e}al, Montr\'{e}al, Qu\'{e}bec, Canada

$^*$ To whom correspondence should be addressed:
\url{will.pearse@usu.edu} and \url{j.davies@ubc.ca}

\textbf{Word-count}: XXX (excluding figure legends)

\clearpage
\textbf{Abstract. Here we
  show ... Abstract.}

% Species have climatic tolerances that are conserved
% * Caused by deep-time evolutionary processes we can detect
% * Tested by local-scale ecological assembly
% * Shown by niche-models
Species' distributions are determined by interacting factors such as
traits, climate, land-use, biotic interactions, and dispersal
limitation \supercite{Pulliam2000,Ricklefs2004,Vellend2010}. Unpicking
the relative contributions and mechanisms of these processes is
challenging, but critical if we are to make predictions about
biodiversity as these factors change simultaneously in the
Anthropocene \supercite{Heller2009,Oliver2014,Urban2015}. While
macroevolutionary biologists frequently argue that species'
distributions tend to shown phylogenetic conservatism, the tendency
for closely-related species to have similar niches
\supercite{Wiens2004,Wiens2010}. Likewise, ecologists have shown that
species' distributions are strongly driven by climate, either through
correlative studies \supercite{xxx} or experimental studies
\supercite{xxx}. While environment and climate are clearly not the
only determinants of species' distributions, decades of research into
niche distribution models have upheld the view that they are important
\supercite{xxx}.

% Species are responding to climate change
% * Species' present-day distribution still show phylogenetic signal
% * Species show local-scale assembly changes in manipulations and long-term studies
% * Ecological niche modelers show that distributions are changing
Species' distributions are undoubtedly responding to climate
\supercite{Parmesan2003,Parmesan2006} and land-use
\supercite{Newbold2015} changes. Species' contemporary distributions
still, however, exhibit phylogenetic signal \supercite{xxx}, implying
that these changes are continuations of past conditions. Equally,
experimental studies \supercite{xxx}, and the ever-increasing number
of observational studies of ongoing climate change \supercite{xxx},
have shown that species' distributions and local-scale ecological
assembly respond to climate. Niche modelers have shown that their
models have some predictive power over species' distributions given
past relationships with climate \supercite{xxx}. That there is broad
consensus that ranges are changing does not mean there have not been
debates over the precise implications and mechanisms of such
changes. For example, the implications of range change for local-scale
biodiversity is unclear
\supercite{Vellend2013,Dornelas2014,Gonzalez2016}, we do not know
whether species' environmental \emph{limits} or \emph{central
  tendencies} are most-strongly conserved
\supercite{deCasas2017,Zanne2018}, and there is a relative paucity of
historic data with which to create historic
baselines\supercite{Newbold2015,Meineke2018}.

% Yet we do not know the extent to which species are tracking their niches
% * We know there is idiosyncracy and that niche models do not always work well
% * Could this be because we have yet to define the null model by which they should be assembling?
% * Could this be because of a breakdown in the climate vectors of species? That heterogeneity is causing problems for species to track their climate?
% * Here we present a framework that accounts for this.
This consensus of species' past and present ranges broadly tracking
the same underlying climate optima is complicated by paleo-ecological
studies that have revealed greater variability of species' historic
ranges in previous centuries and millennia than previously thought
\supercite{Williams2007,Veloz2012,Maguire2015}. Critically, while
species' present-day distributions have changed, and they are broadly
similar to past distributions, we have yet to establish a historic
benchmark by which to assess the magnitude of that tracking. Many
species' distributions span continents, while others spans kilometers,
and the velocity of climate change is not constant across those scales
\supercite{xxx}. As climate continues to change, the underlying
associations between axes of climate have changed \supercite{xxx}: if
species are, as they must be \supercite{xxx}, ultimately limited by a
single environmental axis, what are the consequences of this change
for species' distributions? We argue there is a need for a synthetic
measure of the extent to which species have tracked their niche
through time, given the changes to which they have been exposed.

\subsection*{Results and discussion}
\textbf{A comparative index of species' niche tracking.} The efficacy
of niche models to predict species' distributions is, in part, a
measure of the extent to which species are tracking climate. Here we
present a simple metric of species' degree of track, which we define
as:

\begin{equation}
  track_\tau = \frac{present_\tau - past_\tau}{projected_\tau - past_\tau}
\end{equation}

Where $current_\tau$ is a given quantile ($\tau$) of a species'
climatic range in the present, $past_\tau$ the same quantile of a
species' range in the present, and $projected_\tau$ the same quantile
of a species' present-day range under past climatic conditions. Thus a
$track$ of 0 would indicate a species that has tracked changing
climate perfectly, while a $track$ of 1 would be consistent with a
species that had not moved whatsoever to track changing environmental
conditions. Values between 0 and 1 suggest moderate tracking, while
values greater than 1 or less than 0 represent a change in climate
relationship greater than if the species had not moved at
all. Uncertainty in either species' past or present ranges can be
accounted for through a bootstrap procedure, whereby the past and/or
present distributions are re-sampled (with replacement) to match the
sampling of the other time-period, and the standard deviation of the
index across those samples reported.

Our approach has two main innovations: (1) a focus on the whole of a
species' distribution and on (2) expected change. (1) We suggest that
something of a disconnect is developing among related fields of
ecology: species' range modelers typically focus on species'
limits\supercite{xxx} (\emph{i.e.}, the upper and lower quantiles)
while community ecologists focus on central tendencies\supercite{xxx}
(\emph{i.e.}, the median). While there is some evidence plants' global
distributions are more strongly driven by limits
\supercite{Zanne2018}, we see no strong \emph{a priori} reason that
species can, or must, track all quantiles equally\supercite{xxx}. Thus
our approach is to check all quantiles of a species' distribution. (2)
The denominator of our index re-scales species' change around an
expectation under distributional stasis. We suggest that species'
changes should be measured on the basis of the magnitude of change to
which they are exposed. This change reflects the magnitude of
ecological and evolutionary selection, and while we suggest there are
many ways of addressing this, we suggest ours is the simplest and so
perhaps a fair starting point.

\textbf{Niche tracking across the tree of life throughout the past
  sixty years.} To assess the extent to which species have tracked
climate, we used the most comprehensively sampled time-series data for
species' distributions (GBIF\supercite{xxx}) and climate (UEA's
CRU\supercite{xxx}). We focus on only five quantiles within the data:
the \nth{5}, \nth{25}, \nth{50} (median), \nth{75}, and \nth{95}. In
the methods we give more detail of our reproducible analysis pipeline,
and in the supplementary materials we repeat our analyses with more,
and less, stringent data curation and cleaning. Here, however, we
present species' range data collected from 1955--2015 for XXX species
(XXX plants, XXX fungi, XXX mammals, XXX birds, XXX reptiles, XXX
amphibians, and XXX insects), each observation matched to one of nine
climate variables in the year in which they were sampled (mean,
minimum, maximum temperature, diurnal temperature range,
precipitation, wet-day frequency, vapor pressure, and cloud
cover). We present, in the methods and supplementary materials, a more
detailed analysis of how we accounted for potential and
well-known\supercite{xxx} biases in these data, and we emphasize that
our metric was designed explicitly to account for the known increase
in sampling effort through time in GBIF data. We also highlight that
our data support known linkages between species' climatic ranges and
phylogenetic history and traits, suggesting our data reveal true
underlying signal.

Figure \ref{violin} shows the empirical index values for our focal
taxa for mean temperature. Clear patterns emerge: a large number of
species are, to some extent, tracking climate through time (XXX\%
track at least one quantile, XXX\% track at least three, and XXX\%
track all five). Of these taxa, XXX are, on average, tracking the best
(XXX\% track at least one quantile) while XXX track the worst (XXX\%
track at least one quantile).
% To discuss: these percentages go up if we use the absolutised
% version of the index, but I'm not sure that's appropriate anymore.
Results for the other eight climate variables are comparable, and are
presented in the supplementary materials. Notably, however, many taxa
are not tracking climate through time, and many have experienced more
change than if their distributions had remained static (XXX\% have
`overshot' in at least three quantiles). These results are consistent
with null simulations we performed of species' ranges that included
some degree of climate tracking and empirically observed amounts of
sampling effort changes through time and space in the GBIF data.

\textbf{Idiosyncratic drivers of species' tracking.} To examine the
potential drivers of species' ranges and range tracking, we first
examined the phylogenetic signal of species' climate quantiles. We
focused on mammals\supercite{xxx}, birds\supercite{xxx},
amphibians\supercite{xxx}, reptiles\supercite{xxx}, and
plants\supercite{xxx}, since we were able to find broadly inclusive,
dated phylogenies for these taxa. We found statistically significant
evidence of phylogenetic signal across XXX of 135
taxon-climate-quantile combinations of species' past ranges (and XXX
in the present; see figure \ref{phylogeny}). This is consistent with
the known patterns of phylogenetic conservatism outlined above. For
our index, however, we found evidence of phylogenetic signal in only
XXX combinations (XXXlist themXXX). This confirms the known results
that species' present-day and past ranges are strongly influenced by
evolutionary, but suggests that the magnitude of change (relative to
expectations) may be driven by different drivers.

We additionally examined species' traits as potential drivers of
species' distributions and potential to track changing environment. We
focused on mammals\supercite{xxx}, birds\supercite{xxx}, and
plants\supercite{xxx}, since global, openly-released trait datasets
exist for these taxa. Again, we found statistically significant
associations between traits and past distributions of XXX of XXX
taxon-climate-quantile combinations (XXX in the present; see figure
\ref{traits}). While the traits we used are strongly associated with
species' climate responses\supercite{xxx} (hence their strong
predictive power of species' distributions, but not tracking index),
we emphasize that we do not examine traits related to dispersal
ability. This is due to a paucity of open-access dispersal traits
\supercite{xxx}, and we highlight that future work should address
whether our index is explained by such drivers.
% Another potential problem here is no trait*env interaction, which I
% sense is also important...
% WILL - re-do TEMP.INDEX ~ TEMP! THERMOPHILES!!!

\textbf{Climate space is not the same as tracking space.} Given our
index is poorly predicted by traditionally important variables, we now
consider whether index values are predictive of each other. It is well
known that global climate\supercite{xxx}, and so species' positions
across the various axes of climate space\supercite{xxx}, are not
independent, and so we set out to test whether species' tracking
indices are also independent. As we show in figure \ref{pca}, a
principal components analysis of species' present, past, and projected
climate distributions shows strong correlations across climate
variables. Yet the index does not: species' relative degree of
tracking is both of much higher dimension, and shows relatively less
correlation across climates axes. We argue, therefore, that while
species' positions in climate space show strong associations, species'
relative tracking of their position in climate space does not.

If true, this result would explain the lack of association between
species' traits and evolutionary history with their degree of climate
tracking. The ecological and evolutionary rules that determine
species' climate relationships are changing, and thus the factors that
once are strong predictors of past and present distributions are not
predictive of the differences between them. Yet there are alternative
explanations, of which we highlight three: (1) that we have missed
fundamentally important environmental axes, (2) biotic interactions,
and (3) that our index highlights random variation in
distribution. (1) We fully acknowledge that we ignore potentially
relevant environmental drivers such as geology (and so soil), habitat,
and all Anthropogenic factors other than climate. We are not alone in
this simplification\supercite{xxx}, but we detect strong evolutionary
pattern in the factors we do examine in the data underlying our
index. (2) Competition and facilitation among species are likely
strong drivers of species' ability to track climate\supercite{xxx}. We
acknowledge it is entirely possible that ignoring these factors is the
driver of our result and, if so, we urge for an acceleration of the
inclusion of such information into broad-scale analyses
\supercite{xxx}. (3) If species were perfectly tracking climate
(\emph{i.e.}, $present_\tau$ is equal to $past_\tau$ in the absence of
sampling error or some other stochastic process) then our index values
would essentially be random numbers. There are good empirical evidence
that this is not the case \supercite{xxx} but, as we show in the
supplementary materials, simulations suggest out values are
incompatible with this result.

\subsection*{Conclusion}
Our results are consistent with the processes underlying species'
responses to climate change being distinct from those that generated
their ranges in the past. That is not to say that we do not detect the
imprint of ecological and evolutionary history, or that those earlier
processes are of predictive power for species distributions, but
rather that to understand change we must consider new drivers. We do
not know what those drivers are, although we highlight that biotic
interactions may play a part. We also highlight that our data,
although comprehensive, are not taxonomically complete (we address
only XXX species) and are necessarily spatially ($0.5^\circ$ grid
cells) and temporally (yearly) coarse given the extent (global and 60
years) of our study. We are not the first to suggest that, for large
numbers of species, the extent of climate tracking is idiosyncratic
\supercite{xxx}. These results are, however, consistent with the
paleo-ecological record. During and after the last ice-age,
`no-analogue' communities generated previously-unseen species
assemblages \supercite{xxx}. Once previous mass extinction events were
underway, species' extinctions became unpredictable
\supercite{xxx}. Scientists have been warning for decades that we are
beginning a sixth mass extinction;
% Soule 1999 but find something earlier
perhaps we are seeing the end of its beginning.

\clearpage
% Figures

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.7\textwidth]{../figures/violin-tmp-50.pdf}
  \end{center}
  \caption{\doublespacing \textbf{Empirical distributions of track
      index for \nth{50} of mean temperature.} Need to add simulations
    to plot. Add phylo-pics?}
  \label{pca}
\end{figure}

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.7\textwidth]{../figures/pca-50.pdf}
  \end{center}
  \caption{\doublespacing \textbf{Principal components analyses of
      present and past climate, along with our index of climate
      tracking.} Loadings are correlated (eigenvalues shown at bottom
    of plot), but much weaker overall in index.}
  \label{pca}
\end{figure}

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.7\textwidth]{../figures/phylo-tmp-50.pdf}
  \end{center}
  \caption{\doublespacing \textbf{Phylogeny with ancestral states
      branches of past temperature (50th quantile) and tips of the
      temperature 50th index.} Scale between the two kinds of input
    data currently is slightly off, hence no scale bar, and obviously
    the tip colours are hideous, but I think no one would disagree
    there's signal in the underlying data!}
  \label{pca}
\end{figure}

% zoom in on one clade and show that the tips are a random subset (but the branches are not)
% Will wonders if we can put the correlation matrix to the side of an upward-facing tree
% have lines going outward from the tree to show mapping

\begin{figure}[h!]
  \begin{center}
  \end{center}
  \caption{\doublespacing \textbf{Trait correlation box.}
    XXXmissingXXX}
  \label{traits}
\end{figure}

\clearpage

\textbf{Acknowledgments} WDP was funded by NSF ABI-1759965, NSF
EF-1802605, and USDA Forest Service agreement 18-CS-11046000-041. TJD
was funded by \emph{Fonds de Recherche Nature et Technologies} grant
number 168004.

\textbf{Author Contributions} WDP and TJD contributed to all aspects
of the study.

\textbf{Author information} Reprints and permissions information is
available at \url{www.nature.com/reprints}. The authors declare no
competing financial interests. Correspondence and requests for
materials should be addressed to \url{will.pearse@usu.edu}.

\textbf{Competing interests} The authors declare no competing
financial interests

\clearpage

\section*{\Large Methods}

\printbibliography

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2019-10-22 src/analysis-traits.R

@@ -11,63 +11,90 @@ source("src/headers.R")
     colnames(output) <- names(data); rownames(output) <- c("r","p")
     return(output)
 }
+.combine.data <- function(index, quant, clean=10, abs=FALSE){
+    data <- .simplify.group(index, quant, clean, abs)
+    data$species <- rownames(data)
 
-# Load data
+    data$bm.neotoma <- neotoma$log.mass[match(data$species, neotoma$binomial)]
+    data$bm.elton <- elton.bm[data$species]
+
+    data$log.ll <- glopnet$log.LL[match(data$species, glopnet$Species)]
+    data$log.lma <- glopnet$log.LMA[match(data$species, glopnet$Species)]
+    data$log.Nmass <- glopnet$log.Nmass[match(data$species, glopnet$Species)]
+    data$log.Amass <- glopnet$log.Amass[match(data$species, glopnet$Species)]
+    return(data)    
+}
+.corr.mat <- function(index, quant, clean, abs=FALSE){
+    data <- .combine.data(index, quant, clean, abs)
+
+    c.var <- setNames(
+        c("cloud-cover","frost-days","potential-evapotranspiration","precipitation","min-temperature","mean-temperature","max-temperature","vapour-pressure"),
+                      c("cld","frs","pet","pre","tmn","tmp","tmx","vap")
+    )
+    q.var <- setNames(
+        c("5th","25th","50th","75th","95th"),
+        c("0.05","0.25","0.5","0.75","0.95")
+    )[quant]
+    t.var <- setNames(
+        c("body-mass","body-mass","leaf-lifespan","leaf-mass/area","leaf-N","photosynthetic-capacity"),
+        c("bm.neotoma","bm.elton","log.ll","log.lma","log.Nmass","log.Amass")
+    )
+    
+    mat <- matrix(nrow=6, ncol=8, dimnames=list(t.var, c.var))
+    for(i in seq_along(c.var)){
+        for(j in seq_along(t.var)){
+            mat[j,i] <- cor.test(data[,names(c.var)[i]], data[,names(t.var)[j]])$estimate
+        }
+    }
+    return(mat)
+}
+.plot.corr.mat <- function(index, comparison, quant, clean, abs=FALSE){
+    index.mat <- t(.corr.mat(index, quant, clean, abs))
[…]

### 2019-10-22 src/analysis-violin.R

@@ -2,6 +2,7 @@
 source("src/headers.R")
 
 # Load data
+
 plants <- readRDS("clean-data/plants-index.RDS")
 fungi <- readRDS("clean-data/plants-index.RDS")
 insects <- readRDS("clean-data/plants-index.RDS")
@@ -9,33 +10,56 @@ mammals <- readRDS("clean-data/mammals-index.RDS")
 reptiles <- readRDS("clean-data/reptiles-index.RDS")
 birds <- readRDS("clean-data/birds-index.RDS")
 
-# Plots
-pdf("figures/violin-tmp-50.pdf")
-index <- .simplify.group("index", quantile="0.5", clean=10, abs=FALSE)
-se <- .simplify.group("mad.index", quantile="0.5", abs=FALSE)
-se <- se[se$tmp < 5,]
-se <- se[rownames(se) %in% rownames(index),]
-index <- index[rownames(index) %in% rownames(se),]
-data <- data.frame(index=index$tmp, se=se$tmp, taxon=index$taxon)
-cols <- setNames(c("forestgreen", "black", "blue", "orange", "red", "grey60"), unique(index$taxon))
-data$col <- alpha(cols[data$taxon], alpha=1-((data$se^2)/max(data$se^2)))
-data <- data[order(data$taxon, data$index),]
-data$x <- as.numeric(factor(data$taxon))
-counts <- table(data$taxon)
-curr <- 0
-for(i in seq_along(counts)){
-    taxon <- names(counts)[i]
-    data$x[data$taxon==taxon] <- data$x[data$taxon==taxon] + seq(-.3, .3, length.out=counts[i])
+# Wrapper functions
+.violin.plot <- function(climate, quant, clean=10, se.thresh=5, abs=FALSE, index="bootstrap.index"){
+    # Load and match data
+    index <- .simplify.group(index, quantile=quant, clean=clean, abs=abs)
+    se <- .simplify.group("mad.index", quantile=quant, abs=abs)
+    se <- se[se[,climate] < se.thresh,]
+    se <- se[rownames(se) %in% rownames(index),]
+    index <- index[rownames(index) %in% rownames(se),]
+    data <- data.frame(index=index[,climate], se=se[,climate], taxon=index$taxon)
+
+    # Prepare colours and order for plotting
+    cols <- setNames(c("forestgreen", "black", "blue", "orange", "red", "grey60"), unique(index$taxon))
[…]

### 2019-10-22 src/calculate-index.R

@@ -2,7 +2,7 @@
 source("src/headers.R")
 
 # Wrapper functions
-.calc.track.parallel <- function(data, quantiles, rasters){
+.calc.track.parallel <- function(data, quantiles, rasters, taxon){
     .int.calc <- function(sp){
         # Setup and subset data
         curr.data <- data[scientificname == sp,]
@@ -14,7 +14,7 @@ source("src/headers.R")
             dimnames=list(names(rasters), as.character(quantiles), c("index","bootstrap.index", "mad.index", "quantile","present","past","projected"))
         )
         if(nrow(pre.data) == 0 | nrow(post.data) == 0)
-            return(output)        
+            return(FALSE)        
         
         # Loop over rasters
         for(i in seq_along(rasters)){
@@ -36,12 +36,12 @@ source("src/headers.R")
             output[i,,] <- t(sapply(quantiles, function(x) .calc.metric(pre.data$pre.env, post.data$env, pre.data$post.env, x)))
         }
 
-        return(output)
+        saveRDS(output, paste0(output.dir,taxon,"/",sp,".RDS"))
+        return(TRUE)
     }
-
-    output <- mcMap(.int.calc, unique(data$scientificname))
-    output <- abind(output, along=4)
-    return(output)
+    
+    mcMap(.int.calc, unique(data$scientificname))
+    return(TRUE)
 }
 
 
@@ -71,10 +71,17 @@ cru <- list(
 )    
     
 # Do work and save results
-saveRDS(.calc.track.parallel(plants, quantiles, cru), paste0(output.dir,"index-plants.RDS"))
-saveRDS(.calc.track.parallel(mammals, quantiles, cru), paste0(output.dir,"index-mammals.RDS"))
-saveRDS(.calc.track.parallel(amphibians, quantiles, cru), paste0(output.dir,"index-amphibians.RDS"))
-saveRDS(.calc.track.parallel(fungi, quantiles, cru), paste0(output.dir,"index-fungi.RDS"))
-saveRDS(.calc.track.parallel(birds, quantiles, cru), paste0(output.dir,"index-birds.RDS"))
-saveRDS(.calc.track.parallel(insects, quantiles, cru), paste0(output.dir,"index-insects.RDS"))
-saveRDS(.calc.track.parallel(reptiles, quantiles, cru), paste0(output.dir,"index-reptiles.RDS"))
+print("mammals")
[…]

### 2019-10-22 src/headers.R

@@ -27,7 +27,10 @@ if(!require("scales")){
     install.packages("scales"); library(scales)}
 if(!require("reldist")){
     install.packages("reldist"); library(reldist)}
-
+if(!require("gdata")){
+    install.packages("gdata"); library(gdata)}
+if(!require("corrplot")){
+    install.packages("corrplot"); library(corrplot)}
 
 ####################################
 # Global settings ##################
@@ -199,3 +202,517 @@ prog.bar <- function(x, y){
         data.frame(.simplify(reptiles, index, quantile, clean, abs), taxon="reptiles"),
         data.frame(.simplify(birds, index, quantile, clean, abs), taxon="birds")
     ))
+
+
+
+will.corrplot <- function (corr, method = c("circle", "square", "ellipse", "number", 
+    "shade", "color", "pie"), type = c("full", "lower", "upper"), 
+    add = FALSE, col = NULL, bg = "white", title = "", is.corr = TRUE, 
+    diag = TRUE, outline = FALSE, mar = c(0, 0, 0, 0), addgrid.col = NULL, 
+    addCoef.col = NULL, addCoefasPercent = FALSE, order = c("original", 
+        "AOE", "FPC", "hclust", "alphabet"), hclust.method = c("complete", 
+        "ward", "ward.D", "ward.D2", "single", "average", "mcquitty", 
+        "median", "centroid"), addrect = NULL, rect.col = "black", 
+    rect.lwd = 2, tl.pos = NULL, tl.cex = 1, tl.col = "red", 
+    tl.offset = 0.4, tl.srt = 90, cl.pos = NULL, cl.lim = NULL, 
+    cl.length = NULL, cl.cex = 0.8, cl.ratio = 0.15, cl.align.text = "c", 
+    cl.offset = 0.5, number.cex = 1, number.font = 2, number.digits = NULL, 
+    addshade = c("negative", "positive", "all"), shade.lwd = 1, 
+    shade.col = "white", p.mat = NULL, sig.level = 0.05, insig = c("pch", 
+        "p-value", "blank", "n", "label_sig"), pch = 4, pch.col = "black", 
+    pch.cex = 3, plotCI = c("n", "square", "circle", "rect"), 
+    lowCI.mat = NULL, uppCI.mat = NULL, na.label = "?", na.label.col = "black", 
+    win.asp = 1, ...) 
+{
[…]

### 2019-10-22 ms/ms.tex
Track stats; intro

@@
 \section*{Title page}
-\textbf{Article title}: Placing species' climate envelope within its
-historic context reveals remarkable idiosyncrasy of distributional
-change
+\textbf{Article title}: The relative velocity of climate change
+reveals remarkable idiosyncrasy of species' distributional shifts
 
@@
 \clearpage
-\textbf{Abstract. Here we
-  show ... Abstract.}
+\textbf{To date, our understanding of how species have shifted in
+  response to recent climate warming has been based on a few studies
+  with limited number of species. Here we present a comprehensive,
+  global overview of species' distributional responses to changing
+  climate across a broad variety of taxa (animals, plants, and
+  fungi). We characterise species' responses using a metric that
+  describes the biotic velocity of climate change: how closely
+  species' responses have tracked changing climate through time
+  climate. In contrast to existing metrics that focus purely on space,
+  we focus on species, and thus can examine drivers of interspecific
+  variation. Here we show that species responses are highly
+  idiosyncratic, and thus previous analyses assuming uniform responses
+  may be misleading. This is in stark contrast to species' static
+  present-day and past distributions, which show strong evidence of
+  the imprint of evolutionary history and functional traits. Our
+  analyses are a first step in exploring the vast wealth of empirical
+  data on species' historic responses to recent climate change. We
+  highlight that our findings, while surprising, are consistent with
+  paleontological patterns observed at the onset of mass extinction
+  events.}
+% Is that last sentence a bit much?
 
-% Species have climatic tolerances that are conserved
-% * Caused by deep-time evolutionary processes we can detect
-% * Tested by local-scale ecological assembly
-% * Shown by niche-models
-Species' distributions are determined by interacting factors such as
-traits, climate, land-use, biotic interactions, and dispersal
-limitation \supercite{Pulliam2000,Ricklefs2004,Vellend2010}. Unpicking
-the relative contributions and mechanisms of these processes is
-challenging, but critical if we are to make predictions about
-biodiversity as these factors change simultaneously in the
-Anthropocene \supercite{Heller2009,Oliver2014,Urban2015}. While
-macroevolutionary biologists frequently argue that species'
-distributions tend to shown phylogenetic conservatism, the tendency
-for closely-related species to have similar niches
-\supercite{Wiens2004,Wiens2010}. Likewise, ecologists have shown that
-species' distributions are strongly driven by climate, either through
-correlative studies \supercite{xxx} or experimental studies
-\supercite{xxx}. While environment and climate are clearly not the
-only determinants of species' distributions, decades of research into
-niche distribution models have upheld the view that they are important
-\supercite{xxx}.
+The natural environment is changing rapidly. Over the last century,
+habitats have become increasingly fragmented and
+isolated\supercite{Haddad2015}, the climate has become warmer, and
+extreme climate events more frequent\supercite{IPCC2014}. Biological
+diversity---the species of plants and animals with which we cohabit
+the Earth---has responded predictably. Present-day extinction rates
+are estimated to be three orders of magnitude greater than background
+rates \supercite{Pimm2014} and are projected to increase further over
+the next several decades \supercite{MEA2005}. Current estimates
+suggest that over at least one million species may be threatened with
+extinction\supercite{IPBESland2019}, and that we have experienced a
+50\% decline in animal diversity over the past 40 years
+\supercite{McLellan2014}.
 
-% Species are responding to climate change
-% * Species' present-day distribution still show phylogenetic signal
-% * Species show local-scale assembly changes in manipulations and long-term studies
-% * Ecological niche modelers show that distributions are changing
-Species' distributions are undoubtedly responding to climate
-\supercite{Parmesan2003,Parmesan2006} and land-use
-\supercite{Newbold2015} changes. Species' contemporary distributions
-still, however, exhibit phylogenetic signal \supercite{xxx}, implying
-that these changes are continuations of past conditions. Equally,
-experimental studies \supercite{xxx}, and the ever-increasing number
-of observational studies of ongoing climate change \supercite{xxx},
-have shown that species' distributions and local-scale ecological
-assembly respond to climate. Niche modelers have shown that their
-models have some predictive power over species' distributions given
-past relationships with climate \supercite{xxx}. That there is broad
-consensus that ranges are changing does not mean there have not been
-debates over the precise implications and mechanisms of such
-changes. For example, the implications of range change for local-scale
-biodiversity is unclear
-\supercite{Vellend2013,Dornelas2014,Gonzalez2016}, we do not know
-whether species' environmental \emph{limits} or \emph{central
-  tendencies} are most-strongly conserved
-\supercite{deCasas2017,Zanne2018}, and there is a relative paucity of
-historic data with which to create historic
-baselines\supercite{Newbold2015,Meineke2018}.
+The main direct human-induced drivers that impact biodiversity now are
+habitat loss \supercite{Newbold2015} and fragmentation
+\supercite{Haddad2015}, but climate change is likely to become a
+dominant driver in the next few
+decades\supercite{Parmesan2003,Thuiller2007,Urban2015}. Projected
+impacts of climate change on biodiversity have attracted much
+attention, and it has been suggested that up to 37\% of species may be
+committed to global extinction by 2050\supercite{Thomas2004}. In the
+face of changing climate species must either shift in space, to track
+favourable conditions, or time, for example, flowering and breeding
+earlier, or migrating sooner. Evidence suggests they are doing both
+\supercite{Parmesan2003,Parmesan2006,Menzel2006,Chen2011}, but not all
+species are responding equally. Some species are expanding their
+distributions, sometimes invasively \supercite{Hellmann2008}, and some
+species are flowering later\supercite{Cook2012}; species may also be
+simply becoming less predictable \supercite{Pearse2017phenology}.
+% I feel this isn't just a token cite; we want to set up our result a
+% little, I think
+Developing robust predictions of ecological responses to climate
+change thus requires data from diverse species, ecosystems, and
+climates to account for these varying drivers. By contrast, current
+projections frequently assume a simplistic or uniform response, and
+have tended to overlook interspecific variation.
+% I'm not sure I necessarily agree with this statement
 
-% Yet we do not know the extent to which species are tracking their niches
-% * We know there is idiosyncracy and that niche models do not always work well
-% * Could this be because we have yet to define the null model by which they should be assembling?
-% * Could this be because of a breakdown in the climate vectors of species? That heterogeneity is causing problems for species to track their climate?
-% * Here we present a framework that accounts for this.
-This consensus of species' past and present ranges broadly tracking
-the same underlying climate optima is complicated by paleo-ecological
-studies that have revealed greater variability of species' historic
-ranges in previous centuries and millennia than previously thought
-\supercite{Williams2007,Veloz2012,Maguire2015}. Critically, while
-species' present-day distributions have changed, and they are broadly
-similar to past distributions, we have yet to establish a historic
-benchmark by which to assess the magnitude of that tracking. Many
-species' distributions span continents, while others spans kilometers,
-and the velocity of climate change is not constant across those scales
-\supercite{xxx}. As climate continues to change, the underlying
-associations between axes of climate have changed \supercite{xxx}: if
-species are, as they must be \supercite{xxx}, ultimately limited by a
-single environmental axis, what are the consequences of this change
-for species' distributions? We argue there is a need for a synthetic
-measure of the extent to which species have tracked their niche
-through time, given the changes to which they have been exposed.
+We currently lack robust estimates of how species are shifting their
+distributions in response to recent warming. The velocity of
+temperature change---the rate a species would have to shift its range
+to maintain constant temperature---has been calculated at 0.42 km/yr
+\supercite{Loarie2009}. A meta-analysis of taxonomic groupsestimated
+the average pole-ward shift in species' distributions has been
+estimated as approximately 1.69 km/yr-1 \supercite{Chen2011},
+suggesting that species might be capable of keeping pace with shifting
+climate. Yet this estimate is derived from just a handful of studies,
+and does not address the large variation in response among species.
+Here we leverage the vast volume of distributional data stored in
+digital data repositories\supercite{url_gbif} to provide the first
+global synthesis of species distributional responses to climate
+change. By contrasting historical niche dimensions with those of the
+present day, we can characterise the true biotic velocity of climate
+change, which represents the actual velocity with which species track
+climate across space.
 
+We show that, on average, species' historical (pre-1970) ranges have
+warmed significantly, and that species have generally shifted their
+distributions to ameliorate this warming. Yet this general pattern
+masks a remarkably degree of variability in species responses: many
+species' ranges are now even warmer than if their distibutions had
+remained static. Further, we find that, while species' distributions
+are strongly predicted by their evolutionary history and functional
+traits, the same is not true of their degree of tracking. Thus while
+species' static distributions are reasonably predictable, their
+changes in distributions are not.
+
+\clearpage
 \subsection*{Results and discussion}
-\textbf{A comparative index of species' niche tracking.} The efficacy
-of niche models to predict species' distributions is, in part, a
-measure of the extent to which species are tracking climate. Here we
-present a simple metric of species' degree of track, which we define
-as:
+\textbf{A comparative index of species' niche tracking.} We propose to
+measure the extent to which a species' distribution is tracking
+climate change by standardising its observed change by the change it
+would have experienced had it not shifted its distribution. Formally,
+we define this as:
 
@@
 Where $current_\tau$ is a given quantile ($\tau$) of a species'
-climatic range in the present, $past_\tau$ the same quantile of a
-species' range in the present, and $projected_\tau$ the same quantile
-of a species' present-day range under past climatic conditions. Thus a
-$track$ of 0 would indicate a species that has tracked changing
-climate perfectly, while a $track$ of 1 would be consistent with a
-species that had not moved whatsoever to track changing environmental
-conditions. Values between 0 and 1 suggest moderate tracking, while
-values greater than 1 or less than 0 represent a change in climate
-relationship greater than if the species had not moved at
-all. Uncertainty in either species' past or present ranges can be
-accounted for through a bootstrap procedure, whereby the past and/or
-present distributions are re-sampled (with replacement) to match the
-sampling of the other time-period, and the standard deviation of the
-index across those samples reported.
+climatic distribution across its present-day range, $past_\tau$ the
+same quantile of a species' range in the past, and $projected_\tau$
+the same quantile of a species' present-day range under past climatic
+conditions. This index should be calculated separately for
+temperature, pollution, or whatever environmental variable is of
+interest. Thus a $track$ of 0 would indicate a species that has
+tracked changing climate perfectly, while a $track$ of 1 would be
+consistent with a species that had not moved whatsoever to track
+changing environmental conditions. Values between 0 and 1 suggest
+moderate tracking, while values greater than 1 or less than 0
+represent a change in climate relationship greater than if the species
+had not moved at all. Uncertainty in either species' past or present
+ranges can be accounted for through a bootstrap procedure, whereby the
+past and/or present distributions are re-sampled (with replacement) to
+match the sampling of the other time-period, and the standard
+deviation of the index across those samples reported (see methods).
 
-Our approach has two main innovations: (1) a focus on the whole of a
-species' distribution and on (2) expected change. (1) We suggest that
+Our approach has two main innovations: its focus on (1) the whole of a
+species' distribution and(2) expected change. (1) We suggest that
 something of a disconnect is developing among related fields of
 ecology: species' range modelers typically focus on species'
-limits\supercite{xxx} (\emph{i.e.}, the upper and lower quantiles)
-while community ecologists focus on central tendencies\supercite{xxx}
-(\emph{i.e.}, the median). While there is some evidence plants' global
-distributions are more strongly driven by limits
-\supercite{Zanne2018}, we see no strong \emph{a priori} reason that
-species can, or must, track all quantiles equally\supercite{xxx}. Thus
-our approach is to check all quantiles of a species' distribution. (2)
-The denominator of our index re-scales species' change around an
-expectation under distributional stasis. We suggest that species'
-changes should be measured on the basis of the magnitude of change to
-which they are exposed. This change reflects the magnitude of
-ecological and evolutionary selection, and while we suggest there are
-many ways of addressing this, we suggest ours is the simplest and so
-perhaps a fair starting point.
+limits\supercite{Sexton2009} (\emph{i.e.}, the upper and lower
+quantiles) while community ecologists \supercite{Tilman2004} and
+macro-evolutionary \supercite{Zanne2018} biologists focus on central
+tendencies (\emph{i.e.}, the median). We therefore see no strong
+\emph{a priori} reason that species can, or must, track all quantiles
+equally, and so our approach is to check all quantiles of a species'
+distribution. (2) The denominator of our index re-scales species'
+change around an expectation under distributional stasis. We suggest
+that species' changes should be measured on the basis of the magnitude
+of change to which they are exposed. This change reflects the
+magnitude of ecological and evolutionary selection, and while we
+suggest there are many ways of addressing this, we suggest ours is the
+simplest and so perhaps a fair starting point. We note that our index
+could by re-scaling according to the distance that would need to be
+travelled to match the degree of change (\emph{sensu}
+\citeauthor{Loarie2009}\supercite{Loarie2009}), but prefer to present
+here a dimensionless (and so unitless) index of tracking.
+% ...something like this? I re-wrote everything to be about the
+% "relative velocity of change" (which would have given us some fancy
+% notation if we wanted it) but as we don't use distances or speeds I
+% feel we ought not to?
 
@@
 climate, we used the most comprehensively sampled time-series data for
-species' distributions (GBIF\supercite{xxx}) and climate (UEA's
-CRU\supercite{xxx}). We focus on only five quantiles within the data:
-the \nth{5}, \nth{25}, \nth{50} (median), \nth{75}, and \nth{95}. In
-the methods we give more detail of our reproducible analysis pipeline,
-and in the supplementary materials we repeat our analyses with more,
-and less, stringent data curation and cleaning. Here, however, we
-present species' range data collected from 1955--2015 for XXX species
-(XXX plants, XXX fungi, XXX mammals, XXX birds, XXX reptiles, XXX
-amphibians, and XXX insects), each observation matched to one of nine
-climate variables in the year in which they were sampled (mean,
-minimum, maximum temperature, diurnal temperature range,
-precipitation, wet-day frequency, vapor pressure, and cloud
-cover). We present, in the methods and supplementary materials, a more
-detailed analysis of how we accounted for potential and
-well-known\supercite{xxx} biases in these data, and we emphasize that
-our metric was designed explicitly to account for the known increase
-in sampling effort through time in GBIF data. We also highlight that
-our data support known linkages between species' climatic ranges and
-phylogenetic history and traits, suggesting our data reveal true
-underlying signal.
+species' distributions (GBIF\supercite{url_gbif}) and climate (UEA's
+CRU\supercite{Harris2014}). We focus on five quantiles: the \nth{5},
+\nth{25}, \nth{50} (median), \nth{75}, and \nth{95}. In the methods we
+give more detail of our reproducible analysis pipeline, and in the
+supplementary materials we repeat our analyses with different data
+curation and cleaning choices. Here, however, we present species'
+range data collected from 1955--2015 for 10,700 species (4,879 plants,
+1,517 fungi, 273 mammals, 433 birds, 147 reptiles and amphibians, and
+3,451 insects), each observation matched to one of nine climate
+variables in the year in which they were sampled (mean, minimum,
+maximum temperature, diurnal temperature range, precipitation, wet-day
+frequency, vapour pressure, and cloud cover). Our metric was designed
+to account for known biases in the GBIF data \supercite{Beck2014}, and
+we present in the supplementary materials additional analyses that
+address and account for such biases.
 
 Figure \ref{violin} shows the empirical index values for our focal
-taxa for mean temperature. Clear patterns emerge: a large number of
-species are, to some extent, tracking climate through time (XXX\%
-track at least one quantile, XXX\% track at least three, and XXX\%
-track all five). Of these taxa, XXX are, on average, tracking the best
-(XXX\% track at least one quantile) while XXX track the worst (XXX\%
-track at least one quantile).
-% To discuss: these percentages go up if we use the absolutised
-% version of the index, but I'm not sure that's appropriate anymore.
-Results for the other eight climate variables are comparable, and are
-presented in the supplementary materials. Notably, however, many taxa
-are not tracking climate through time, and many have experienced more
-change than if their distributions had remained static (XXX\% have
-`overshot' in at least three quantiles). These results are consistent
-with null simulations we performed of species' ranges that included
-some degree of climate tracking and empirically observed amounts of
-sampling effort changes through time and space in the GBIF data.
+taxa for mean temperature. Clear patterns emerge: 99.5\% of species
+are tracking some climate variable through time, and a large number of
+species are, to some extent, tracking temperature through time (65\%
+track at least one quantile, 43\% track at least two, and 4\% track
+all five). The three temperature indices are the least-tracked of any
+climate variable, while precipitation is the most-tracked (86\% of
+species track at least one of its quantiles). Notably, however, many
+taxa are not tracking climate through time, and many have experienced
+more change than if their distributions had remained static (99.9\% of
+species have `overshot' in at least three quantiles). As we show in
+figure \ref{violin}, and explore further in the supplementary
+materials, these results are consistent with our null simulations
+where species' range change is a combination of essentially random
+movement and a moderate degree of subsequent climate filtering.
 
@@
 
+This consensus of species' past and present ranges broadly tracking
+the same underlying climate optima is complicated by paleo-ecological
+studies that have revealed greater variability of species' historic
+ranges in previous centuries and millennia than previously thought
+\supercite{Williams2007,Veloz2012,Maguire2015}.
+
 \clearpage
@@
     to plot. Add phylo-pics?}
-  \label{pca}
+  \label{violin}
 \end{figure}

### 2019-10-22 src/analysis-violin.R

@@ -10,7 +10,7 @@ mammals <- readRDS("clean-data/mammals-index.RDS")
 reptiles <- readRDS("clean-data/reptiles-index.RDS")
 birds <- readRDS("clean-data/birds-index.RDS")
 
-# Wrapper functions
+# Do the plots
 .violin.plot <- function(climate, quant, clean=10, se.thresh=5, abs=FALSE, index="bootstrap.index"){
     # Load and match data
     index <- .simplify.group(index, quantile=quant, clean=clean, abs=abs)
@@ -63,3 +63,67 @@ for(climate in c("cld","frs","pet","pre","tmn","tmp","tmx","vap")){
         dev.off()
     }
 }        
+
+# Calculate the numbers
+.percentage.stats <- function(index, var, clean, abs=FALSE, overshoot=FALSE, process=TRUE){
+    index <- .simplify.quantiles(index, var, clean, abs)
+    if(overshoot){
+        index <- index > 1
+    } else index <- index < 1 & index > 0
+    if(process){
+        percentages <- table(rowSums(index)) / nrow(index)
+        percentages <- percentages[order(names(percentages))]
+        something <- 1-percentages[1]
+        at.least.two <- 1 - sum(percentages[1:2])
+        at.least.three <- 1 - sum(percentages[1:3])
+        everything <- percentages[6]
+        return(setNames(
+            c(something,at.least.two,at.least.three, everything),
+            c("something","at.least.two","at.least.three","everything")))
+    } else return(index)
+}
+.percentage.stats("bootstrap.index", "cld")
+.percentage.stats("bootstrap.index", "frs")
+.percentage.stats("bootstrap.index", "pet")
+.percentage.stats("bootstrap.index", "pre")
+.percentage.stats("bootstrap.index", "tmn")
+.percentage.stats("bootstrap.index", "tmp")
+.percentage.stats("bootstrap.index", "tmx")
+.percentage.stats("bootstrap.index", "vap")
+.percentage.stats("bootstrap.index", "wet")
+
+combined <- .percentage.stats("bootstrap.index", "wet", process=FALSE) |
+    .percentage.stats("bootstrap.index", "frs", process=FALSE) |
+    .percentage.stats("bootstrap.index", "pet", process=FALSE) |
+    .percentage.stats("bootstrap.index", "pre", process=FALSE) |
[…]

### 2019-10-26 ms/ms.tex
Complete draft

@@
 \usepackage{amsmath}
+\usepackage{pdflscape}
 
@@
 to maintain constant temperature---has been calculated at 0.42 km/yr
-\supercite{Loarie2009}. A meta-analysis of taxonomic groupsestimated
+\supercite{Loarie2009}. A meta-analysis of taxonomic groups estimated
 the average pole-ward shift in species' distributions has been
@@
 masks a remarkably degree of variability in species responses: many
-species' ranges are now even warmer than if their distibutions had
+species' ranges are now even warmer than if their distributions had
 remained static. Further, we find that, while species' distributions
@@
 something of a disconnect is developing among related fields of
-ecology: species' range modelers typically focus on species'
+ecology: species' range modellers typically focus on species'
 limits\supercite{Sexton2009} (\emph{i.e.}, the upper and lower
@@
 \citeauthor{Loarie2009}\supercite{Loarie2009}), but prefer to present
-here a dimensionless (and so unitless) index of tracking.
+here a dimensionless (and so unit-less) index of tracking.
 % ...something like this? I re-wrote everything to be about the
@@
 examined the phylogenetic signal of species' climate quantiles. We
-focused on mammals\supercite{xxx}, birds\supercite{xxx},
-amphibians\supercite{xxx}, reptiles\supercite{xxx}, and
-plants\supercite{xxx}, since we were able to find broadly inclusive,
-dated phylogenies for these taxa. We found statistically significant
-evidence of phylogenetic signal across XXX of 135
-taxon-climate-quantile combinations of species' past ranges (and XXX
+focused on mammals\supercite{Faurby2015}, birds\supercite{Jetz2012},
+amphibians \supercite{Jetz2018}, reptiles \supercite{Zheng2016}, and
+plants\supercite{Smith2018}, since we were able to find broadly
+inclusive, dated phylogenies for these taxa. We found statistically
+significant evidence of phylogenetic signal across 158 of 180
+taxon-climate-quantile combinations of species' past ranges (and 153
 in the present; see figure \ref{phylogeny}). This is consistent with
@@
 our index, however, we found evidence of phylogenetic signal in only
-XXX combinations (XXXlist themXXX). This confirms the known results
-that species' present-day and past ranges are strongly influenced by
-evolutionary, but suggests that the magnitude of change (relative to
-expectations) may be driven by different drivers.
+29 combinations. This confirms the known results that species'
+present-day and past ranges are strongly influenced by evolutionary,
+but suggests that the magnitude of change (relative to expectations)
+may be driven by different drivers.
+% Will: this is based on lambda *values* and so you need to update
+% this with p-values...
 
@@
 species' distributions and potential to track changing environment. We
-focused on mammals\supercite{xxx}, birds\supercite{xxx}, and
-plants\supercite{xxx}, since global, openly-released trait datasets
-exist for these taxa. Again, we found statistically significant
-associations between traits and past distributions of XXX of XXX
-taxon-climate-quantile combinations (XXX in the present; see figure
-\ref{traits}). While the traits we used are strongly associated with
-species' climate responses\supercite{xxx} (hence their strong
-predictive power of species' distributions, but not tracking index),
-we emphasize that we do not examine traits related to dispersal
-ability. This is due to a paucity of open-access dispersal traits
-\supercite{xxx}, and we highlight that future work should address
-whether our index is explained by such drivers.
+focused on mammals\supercite{Jones2009a,Wilman2014},
+birds\supercite{Wilman2014}, and plants\supercite{Wright2004}, since
+global, openly-released trait datasets exist for these taxa. Again, we
+found statistically significant associations between traits and past
+distributions of 193 of 240 taxon-climate-quantile combinations (203
+in the present at $\alpha_{5\%}$; see figure \ref{traits}), but not
+for between the traits and our tracking index (9 out of 240l
+4\%). While the traits we used are strongly associated with species'
+climate responses\supercite{Wright2004} (hence their strong predictive
+power of species' distributions, but not tracking index), we emphasise
+that we do not examine traits related to dispersal ability. This is
+due to a paucity of open-access dispersal traits
+\supercite{Gallagher2019}, and we highlight that future work should
+address whether our index is explained by such drivers.
 % Another potential problem here is no trait*env interaction, which I
@@
 once are strong predictors of past and present distributions are not
-predictive of the differences between them. Yet there are alternative
+predictive of the differences between them. A likely
+
+Yet there are alternative
 explanations, of which we highlight three: (1) that we have missed
@@
 % Figures
-
+\clearpage
 \begin{figure}[h!]
   \begin{center}
-    \includegraphics[width=0.7\textwidth]{../figures/violin-tmp-50.pdf}
+    \includegraphics[width=\textwidth]{../figures/violin-tmp-0-5.pdf}
   \end{center}
-  \caption{\doublespacing \textbf{Empirical distributions of track
-      index for \nth{50} of mean temperature.} Need to add simulations
-    to plot. Add phylo-pics?}
+  \caption{\doublespacing \textbf{Most species are tracking global
+      temperature, but there is pronounced variation among species and
+      clades.} The vertical axis shows the \nth{50} quantile of our
+    tracking index of annual mean temperature (boostrapped to account
+    for variation in sampling; see methods). On the left, our seven
+    major taxonomic groups plotted with the median and \nth{25} and
+    \nth{50} quantiles of their distributions highlighted. At the
+    far-right of the plot, output from our simulations of species that
+    are moving but either imperfectly tracking, or not tracking at
+    all, are shown. In the main text, we highlight that the majority
+    of species are tracking global climate, but there is profound
+    variation among species and clades. Fungi, insects, and reptiles
+    are, broadly, tracking the \nth{50} quantile of mean temperature
+    well, while other taxa (\emph{e.g.}, mammals and plants) are not.}
   \label{violin}
@@
 
+\clearpage
+\begin{landscape}
 \begin{figure}[h!]
   \begin{center}
-    \includegraphics[width=0.7\textwidth]{../figures/pca-50.pdf}
+    \includegraphics[width=1.4\textwidth]{../figures/phylo-tmp-50.pdf}
   \end{center}
-  \caption{\doublespacing \textbf{Principal components analyses of
-      present and past climate, along with our index of climate
-      tracking.} Loadings are correlated (eigenvalues shown at bottom
-    of plot), but much weaker overall in index.}
+  \caption{\doublespacing \textbf{Despite strong phylogenetic signal
+      in species' static distributions, species' changes in
+      distributions are not phylogenetically patterned.} This is a
+    single sample from the posterior distribution of plant phylogenies
+    used for our analysis\supercite{Smith2018}. The two oldest
+    branches have been shortened for clarity of presentation
+    (emphasised with dashed lines). Its internal branches are
+    ancestral state reconstructions, assuming Brownian motion, of
+    species' median mean-temperature in the past. Its tips are
+    coloured according to the track index of the same underlying data
+    (compare with figure \ref{violin}). Both are coloured according to
+    the legend at the bottom of the figure. While this is but a single
+    plot (see also figure \ref{correls} and the supplementary
+    materials), this general pattern of strong phylogenetic signal of
+    climate, but not of degree of tracking, is true of almost all of
+    our underlying data (see main text).}
   \label{pca}
 \end{figure}
+\end{landscape}
 
+\clearpage
 \begin{figure}[h!]
   \begin{center}
-    \includegraphics[width=0.7\textwidth]{../figures/phylo-tmp-50.pdf}
+    \begin{subfigure}{.45\textwidth}
+      \includegraphics[width=\textwidth]{../figures/traits-50.pdf}
+      \caption{Trait correlations}
+    \end{subfigure}
+    \begin{subfigure}{.45\textwidth}
+      \includegraphics[width=\textwidth]{../figures/physig-5.pdf}
+      \caption{Phylogenetic signal}
+    \end{subfigure}
   \end{center}
-  \caption{\doublespacing \textbf{Phylogeny with ancestral states
-      branches of past temperature (50th quantile) and tips of the
-      temperature 50th index.} Scale between the two kinds of input
-    data currently is slightly off, hence no scale bar, and obviously
-    the tip colours are hideous, but I think no one would disagree
-    there's signal in the underlying data!}
-  \label{pca}
-\end{figure}
-
-% zoom in on one clade and show that the tips are a random subset (but the branches are not)
-% Will wonders if we can put the correlation matrix to the side of an upward-facing tree
-% have lines going outward from the tree to show mapping
-
-\begin{figure}[h!]
-  \begin{center}
-  \end{center}
-  \caption{\doublespacing \textbf{Trait correlation box.}
-    XXXmissingXXX}
+  \caption{\doublespacing \textbf{Species' static climate variables
+      show strong trait correlations and phylogenetic signal, but
+      their relative degree of tracking does not.} In (a), we show the
+    correlation between the median of each climate variable (rows)
+    with each of our traits (columns); colours are given at the right
+    of the figure. In each cell, the outer, larger box depicts the
+    correlation between species' present-day climate and the trait,
+    while the smaller, inner box the correlation for species' relative
+    tracking (whose size is proportional to intensity of
+    correlation). The 3 (of 54) track-index correlations that are
+    statistically significant (at $\alpha_{5\%}$) are highlighted with
+    crosses. In (b), we show the phylogenetic signal (Pagel's
+    $\lambda$\supercite{Pagel1999}) of the median of each taxonomic
+    group's (columns) median species' climate variables and relative
+    tracking indices (rows). As with (a), in each cell the outer,
+    larger box depicts the correlation between species' present-day
+    climate and the trait, while the smaller, inner box the
+    correlation for species' relative tracking (whose size is
+    proportional to phylogenetic signal strength). The single
+    track-index metric for which phylogenetic signal was stronger than
+    the underlying data is highlighted with a cross.}
   \label{traits}
@@
 \clearpage
+\begin{figure}[h!]
+  \begin{center}
+    \includegraphics[width=\textwidth]{../figures/pca-50.pdf}
+  \end{center}
+  \caption{\doublespacing \textbf{Species' movement through climate
+      space differs from their current positions within it.} Each of
+    our figures is a principal component analysis (PCA) biplots; in
+    (a) of species' median past climate values, in (b) median present
+    climate, and in (c) the median track-index of all climate
+    variables. A PCA biplot shows correlations among the underlying
+    data, with each axis explaining some amount of variation in the
+    underlying data (labelled on the axes themselves). In (a) and (b),
+    species' climate variables are correlated with one-another, and
+    the space reflects global patterns in climate. In (c), this
+    pattern has broken down: while some associations among climate
+    axes still remain (notably temperature), the general pattern is
+    one of independent shift through climate space.}
+  \label{pca}
+\end{figure}
 
-\textbf{Acknowledgments} WDP was funded by NSF ABI-1759965, NSF
+\clearpage
+
+\textbf{Acknowledgements} WDP was funded by NSF ABI-1759965, NSF
 EF-1802605, and USDA Forest Service agreement 18-CS-11046000-041. TJD
@@
 \section*{\Large Methods}
+\section*{\Large Methods}
+\renewcommand{\figurename}{Extended Data Figure}
+\renewcommand{\tablename}{Extended Data Table}
+
+All analyses were conducted in \texttt{R} version \texttt{3.6.1}
+\autocite{R2018}. Code to reproduce all analyses in their entirety is
+available in the online supplementary materials.
+
+\subsection*{Data collation}
+
+\textbf{Occurrence data.} Species' distribution data were downloaded
+from the Global Biodiversity Information Facility (GBIF)
+\supercite{url_gbif,url_gbif_occurrences}. We downloaded only data
+highlighted by GBIF as not having any spatial issues, meaning they
+contained no obvious location errors (\emph{e.g.}, latitude/longitude
+swaps, coordinates that are rounded and so imprecise). These occurrence
+data include records that were flagged as having potential issues and
+then were corrected; in the supplementary materials, we show that
+excluding these records leads to qualitatively identical results. We
+present, in the main text, results from `human observation'
+occurrences (\emph{e.g.}, human surveys); in the supplementary
+materials observations based upon vouchered specimens (\emph{e.g.},
+pressed specimens in a herbarium) whose results are qualitatively
+identical. We worked only with occurrences from 1955--1980 and
+1990-2015 (see also `metric calculation' below). We only worked with
+species that had at least 1000 records on GBIF. This pragmatic
+threshold limits us to better-studied species, but in preliminary
+investigations we found our results to be robust to this threshold.
+
+\textbf{Climate data.} Climate data were downloaded from
+\citeauthor{Harris2014}\supercite{Harris2014}'s nine global
+time-series. We split the data into yearly (averaged) $0.5^\circ$
+cells, from which we estimated each of our nine climate variables for
+each species observation.
+
+\textbf{Trait and phylogenetic data.} Plant trait data were taken from
+\citeauthor{Wright2004}\supercite{Wright2004} (leaf lifespan, leaf
+mass-per-unit-area, leaf Nitrogen mass, and leaf photosynthetic
+capacity), mammal body mass data from PanTHERIA\supercite{Jones2009a}
+and EltonTraits \supercite{Wilman2014}, and bird body mass data from
+EltonTraits \supercite{Wilman2014}. We were unable to find sufficient
+coverage of the other taxa in this study in open-access trait
+databases to facilitate further analysis. We used existing global
+mammal\supercite{Faurby2015}, bird\supercite{Jetz2012}, amphibian
+\supercite{Jetz2018}, and plant\supercite{Smith2018} phylogenies,
+which contain posterior distributions of possible phylogenies which we
+used to account for uncertainty (see `Trait and phylogenetic analyses'
+below). For reptiles, we used
+\citeauthor{Zheng2016}\supercite{Zheng2016}, which is a single
+phylogeny and so we could not account for uncertainty with it in this
+way. We were unable to find dated, global phylogenies for the other
+taxa in this study.
+
+\textbf{Taxonomic cleaning.}  We only made use of data for which
+species identity was known, and we ignored variation among
+sub-species. Species' taxonomy has changed through time, and so is a
+potential source of bias in our analysis. As such, we used the World
+Conservation Union's species taxonomy, through
+\emph{taxize}\supercite{Chamberlain2013}, to harmonise all taxonomic
+information. In cases where a species' name could not be found on the
+IUCN, we retained that species in our analysis (since GBIF itself
+performs taxonomic cleaning and standardisation).
+
+\subsection*{Calculating and testing our track index}
+
+\textbf{Bootstrapping procedure to account for sampling bias and
+  empirical calculation of track index values.}  The sampling of
+species through time and across space is known to be uneven across the
+data in GBIF \supercite{Beck2013,Beck2014}. To ensure our method was
+statistically robust to such changes, we employed a bootstrap
+procedure during the calculation of $track_\tau$. To do this we
+randomly re-sampled, with replacement, the occurrences making up
+$present_\tau$ and $projected_\tau$ to be of the same number as
+$past_\tau$, and vice-versa for $past_\tau$ and $present_\tau$. We
+repeated this process 999 times, and report the median of the
+bootstrapped distribution of $track_\tau$ here in the main text. We
+applied this procedure to the occurrence data described above.
+
+\textbf{Estimating the track index in simulated data.} To assess the
+performance of our (bootstrapped) track index, we simulated species'
+whose ranges tracked climate to varying extents through time. Full
+results of the results of these simulations are given in the
+supplementary materials, and a selection of simulations are shown i
+figure \ref{violin}. Using the climate data from above in the years
+1962 and 2002 (the mid-points of the ranges of our data), we simulated
+species with varying maximum possible range sizes (1x1, 2x2, 5x5,
+10x10, 20x20, or 50x50 grid-cell extents) and varying possible
+latitudinal shifts in range ($\pm$1, 2, 5, or 10 grid-cells). We also
+varied two other parameters: a species' degree of environmental
+tracking ($\alpha$; taking a value of 0, 0.1, 0.2, ..., 0.9, or 1) and
+overall fraction of chance occupancy ($\sigma$; taking a value of 0,
+0.1, 0.2, ..., 0.9, or 1). Together, these two parameters define a
+species' probability of being present within a cell within its
+potential range:
+$(1-\alpha) \times sigma + \alpha \times \mathcal{N}$.  $\mathcal{N}$
+is defined as a scaled Normal probability density function with a mean
+equal to the median of the species' distribution in 1962 and a
+variance of 1. Thus a species with an $\alpha$ of 1 is present only in
+cells within its range that resemble its climatic centre, and a
+species with an $\alpha$ of 0 is present in proportion $\sigma$ of the
+cells within its potential range. We simulated all possible
+combinations of parameters 20 times, providing a full range of
+potential range shifts.
+% This needs more work!
+
+\subsection*{Trait and phylogenetic analyses}
+
+We calculated the Pearson's correlation between all estimated climatic
+niche variables ($present_\tau$, $past_\tau$, $projected_\tau$,
+$track_\tau$ and the bootstrapped $track_\tau$). While we show only a
+handful of these correlations in figure \ref{traits}, we provide all
+correlations in the supplementary figures (and note that the
+qualitative results of all correlations are the same). For the
+phylogenetic signal analyses, we likewise calculated Pagel's $\lambda$
+(using \emph{phytools}\supercite{Revell2012phytools}) for all indices,
+reporting all results in full in the supplementary materials (which
+are again qualitatively identical to the results in the main text).
 

### 2019-10-26 src/analysis-pca.R

@@ -19,18 +19,19 @@ source("src/headers.R")
     index <- .simplify.group("bootstrap.index", quantile, clean=clean,  abs=abs)[,1:9]
     present <- .simplify.group("present", quantile, abs=abs)[,1:9]
     past <- .simplify.group("past", quantile, abs=abs)[,1:9]
+    browser()
     
-    layout(matrix(1:6, nrow=2, byrow=TRUE), widths=c(5,5,5), heights=c(5,1))
-    par(mar=c(1,1,1,1))
+    #layout(matrix(1:6, nrow=2, byrow=TRUE), widths=c(5,5,5), heights=c(5,1))
+    par(mar=c(1,1,1,1), mfrow=c(1,3))
     present <- prcomp(present, scale=TRUE)
     past <- prcomp(past, scale=TRUE)
     index <- prcomp(index, scale=TRUE)
     .pca.plot(past, 1:2, bty="o")
     .pca.plot(present, 1:2)
     .pca.plot(index, 1:2)
-    plot(past, main="")
-    plot(past, main="")
-    plot(index, main="")
+    #plot(past, main="")
+    #plot(past, main="")
+    #plot(index, main="")
 }
 
 # Load data

### 2019-10-26 src/analysis-phylo.R

@@ -20,13 +20,14 @@ source("src/headers.R")
         apply(as.matrix(c.data$data), 1, function(x) which.min(abs(x))),
         rownames(c.data$data)
     )
-.phy.sig.wrap <- function(data, tree, method="lambda", merge=TRUE, clean=5){
+.phy.sig.wrap <- function(data, tree, index, clean, method="lambda", merge=TRUE){
     output <- matrix(NA, ncol=9, nrow=5)
     rownames(output) <- dimnames(data)[[2]]
     colnames(output) <- dimnames(data)[[1]]
+    
     for(i in seq_len(nrow(output))){
-        c.data <- .make.comparative.data(.simplify(data, "bootstrap.index", rownames(output)[i], clean=clean), tree, merge)
-        output[i,] <- .phy.sig(c.data)[1,]
+        c.data <- .make.comparative.data(.simplify(data, index, rownames(output)[i]), tree, merge)
+        output[i,] <- as.numeric(.phy.sig(c.data)[1,])
     }
     return(output)
 }
@@ -53,46 +54,156 @@ source("src/headers.R")
     attr(x, "range") <- c(min, max)
     return(x)
 }
+.plot.phylo.mat <- function(quant){
+    q.var <- setNames(
+        c("5th","25th","50th","75th","95th"),
+        c("0.05","0.25","0.5","0.75","0.95")
+    )[quant]
+    c.var <- setNames(
+        c("clouds","frost","evapo-trans.","precipitation","min(temp)","mean(temp)","max(temp)","vapour","rainy day"),
+        c("cld","frs","pet","pre","tmn","tmp","tmx","vap","wet")
+    )
+    
+    index <- matrix(
+        c(readRDS("clean-data/mammals-signal.RDS")$bootstrap.index[quant,],
+          readRDS("clean-data/birds-signal.RDS")$bootstrap.index[quant,],
+          readRDS("clean-data/reptiles-signal.RDS")$bootstrap.index[quant,],
+          readRDS("clean-data/plants-signal.RDS")$bootstrap.index[quant,])
+      , byrow=FALSE, nrow=length(c.var)
+    )
+    comp <- matrix(
+        c(readRDS("clean-data/mammals-signal.RDS")$present[quant,],
[…]

### 2019-10-26 src/analysis-traits.R

@@ -24,11 +24,11 @@ source("src/headers.R")
     data$log.Amass <- glopnet$log.Amass[match(data$species, glopnet$Species)]
     return(data)    
 }
-.corr.mat <- function(index, quant, clean, abs=FALSE){
+.corr.mat <- function(index, quant, clean, abs=FALSE, p.val=FALSE){
     data <- .combine.data(index, quant, clean, abs)
 
     c.var <- setNames(
-        c("cloud-cover","frost-days","potential-evapotranspiration","precipitation","min-temperature","mean-temperature","max-temperature","vapour-pressure"),
+        c("clouds","frost","evapo-trans.","precipitation","min(temp)","mean(temp)","max(temp)","vapour"),
                       c("cld","frs","pet","pre","tmn","tmp","tmx","vap")
     )
     q.var <- setNames(
@@ -36,38 +36,51 @@ source("src/headers.R")
         c("0.05","0.25","0.5","0.75","0.95")
     )[quant]
     t.var <- setNames(
-        c("body-mass","body-mass","leaf-lifespan","leaf-mass/area","leaf-N","photosynthetic-capacity"),
+        c("body-mass","body-mass","leaf lifespan","leaf mass/area","leaf N","photosynth"),
         c("bm.neotoma","bm.elton","log.ll","log.lma","log.Nmass","log.Amass")
     )
     
     mat <- matrix(nrow=6, ncol=8, dimnames=list(t.var, c.var))
     for(i in seq_along(c.var)){
         for(j in seq_along(t.var)){
-            mat[j,i] <- cor.test(data[,names(c.var)[i]], data[,names(t.var)[j]])$estimate
+            if(p.val){
+                mat[j,i] <- cor.test(data[,names(c.var)[i]], data[,names(t.var)[j]])$p.value               
+            } else {
+                mat[j,i] <- cor.test(data[,names(c.var)[i]], data[,names(t.var)[j]])$estimate
+            }
         }
     }
     return(mat)
 }
 .plot.corr.mat <- function(index, comparison, quant, clean, abs=FALSE){
     index.mat <- t(.corr.mat(index, quant, clean, abs))
[…]

### 2019-10-26 src/analysis-violin.R

@@ -2,16 +2,15 @@
 source("src/headers.R")
 
 # Load data
-
 plants <- readRDS("clean-data/plants-index.RDS")
-fungi <- readRDS("clean-data/plants-index.RDS")
-insects <- readRDS("clean-data/plants-index.RDS")
+fungi <- readRDS("clean-data/fungi-index.RDS")
+insects <- readRDS("clean-data/insects-index.RDS")
 mammals <- readRDS("clean-data/mammals-index.RDS")
 reptiles <- readRDS("clean-data/reptiles-index.RDS")
 birds <- readRDS("clean-data/birds-index.RDS")
 
 # Do the plots
-.violin.plot <- function(climate, quant, clean=10, se.thresh=5, abs=FALSE, index="bootstrap.index"){
+.violin.plot <- function(climate, quant, clean=NA, se.thresh=5, abs=FALSE, index="bootstrap.index"){
     # Load and match data
     index <- .simplify.group(index, quantile=quant, clean=clean, abs=abs)
     se <- .simplify.group("mad.index", quantile=quant, abs=abs)
@@ -56,17 +55,69 @@ birds <- readRDS("clean-data/birds-index.RDS")
     axis(1, labels=names(cols), at=1:6)    
     axis(2)
 }
+
+.outline.plot <- function(climate, quant, clean=NA, se.thresh=5, abs=FALSE, index="bootstrap.index"){
+    # Load and match data
+    index <- .simplify.group(index, quantile=quant, clean=clean, abs=abs)
+    se <- .simplify.group("mad.index", quantile=quant, abs=abs)
+    se <- se[se[,climate] < se.thresh,]
+    se <- se[rownames(se) %in% rownames(index),]
+    index <- index[rownames(index) %in% rownames(se),]
+    data <- data.frame(index=index[,climate], se=se[,climate], taxon=index$taxon)
+
+    # Prepare colours and order for plotting
+    cols <- setNames(c("forestgreen", "black", "blue", "orange", "red", "grey60"), unique(index$taxon))
+    data$col <- alpha(cols[data$taxon], alpha=1-((data$se^2)/max(data$se^2)))
+    data <- data[order(data$taxon, data$index),]
+    data$x <- as.numeric(factor(data$taxon))
+    counts <- table(data$taxon)
+    curr <- 0
+    for(i in seq_along(counts)){
+        taxon <- names(counts)[i]
[…]

### 2019-10-26 src/calculate-index.R

@@ -43,7 +43,12 @@ source("src/headers.R")
     mcMap(.int.calc, unique(data$scientificname))
     return(TRUE)
 }
-
+.merge.results <- function(taxon){
+    files <- list.files(paste0(output.dir,taxon))
+    results <- Map(function(x) readRDS(paste0(output.dir,taxon,"/",x)), files)
+    names(results) <- gsub(".RDS", "", names(results), fixed=TRUE)
+    return(abind(results, along=4))
+}
 
 # Load and clean GBIF data
 plants <- .load.gbif(
@@ -85,3 +90,12 @@ print("plants")
 .calc.track.parallel(plants, quantiles, cru, "plants")
 print("birds")
 .calc.track.parallel(birds, quantiles, cru, "birds")
+
+# Merge the results back together again
+saveRDS(.merge.results("mammals"), paste0(output.dir,"mammals-index.RDS"))
+saveRDS(.merge.results("amphibians"), paste0(output.dir,"amphibians-index.RDS"))
+saveRDS(.merge.results("fungi"), paste0(output.dir,"fungi-index.RDS"))
+saveRDS(.merge.results("insects"), paste0(output.dir,"insects-index.RDS"))
+saveRDS(.merge.results("reptiles"), paste0(output.dir,"reptiles-index.RDS"))
+saveRDS(.merge.results("plants"), paste0(output.dir,"plants-index.RDS"))
+saveRDS(.merge.results("birds"), paste0(output.dir,"birds-index.RDS"))

### 2019-10-26 src/headers.R

@@ -31,6 +31,10 @@ if(!require("gdata")){
     install.packages("gdata"); library(gdata)}
 if(!require("corrplot")){
     install.packages("corrplot"); library(corrplot)}
+if(!require("viridis")){
+    install.packages("viridis"); library(viridis)}
+if(!require("plotrix")){
+    install.packages("plotrix"); library(plotrix)}
 
 ####################################
 # Global settings ##################
@@ -181,19 +185,19 @@ prog.bar <- function(x, y){
     }
     return("ERROR")
 }
-.simplify <- function(data, index, quantile, clean, abs=FALSE){
+.simplify <- function(data, index, quantile, clean=NA, abs=FALSE){
     data <- t(data[,quantile,index,])
-    if(!missing(clean)){
+    if(!is.na(clean)){
         keep <- rep(TRUE, nrow(data))
         for(i in seq_len(ncol(data)))
-            keep <- keep & abs(data[,i]) <= clean
+            keep <- keep & abs(data[,i]) <= clean & (!is.na(data[,i]))
         data <- data[keep,]
     }
     if(abs)
         data <- abs(data)
     return(data)
 }
-.simplify.group <- function(index, quantile, clean, abs=FALSE)
+.simplify.group <- function(index, quantile, clean=NA, abs=FALSE)
     return(rbind(
         data.frame(.simplify(plants, index, quantile, clean, abs), taxon="plants"),
         data.frame(.simplify(fungi, index, quantile, clean, abs), taxon="fungi"),
@@ -202,517 +206,12 @@ prog.bar <- function(x, y){
         data.frame(.simplify(reptiles, index, quantile, clean, abs), taxon="reptiles"),
         data.frame(.simplify(birds, index, quantile, clean, abs), taxon="birds")
     ))
-
-
-
-will.corrplot <- function (corr, method = c("circle", "square", "ellipse", "number", 
-    "shade", "color", "pie"), type = c("full", "lower", "upper"), 
-    add = FALSE, col = NULL, bg = "white", title = "", is.corr = TRUE, 
-    diag = TRUE, outline = FALSE, mar = c(0, 0, 0, 0), addgrid.col = NULL, 
-    addCoef.col = NULL, addCoefasPercent = FALSE, order = c("original", 
[…]

### 2019-10-27 ms/ms.tex
Saturday

@@
 \begin{document}
+
+% To-do before bioRxiv:
+% * p-values in lambda
+% * change years in simulation
+% * check taxa numbers in violin plot
+% * add copyright information for phylopics
+% * add sigma into the range simulation code
+% * not a track index. But what is it?...
+% * can you get more birds into the manuscript? They seem to be being dropped somewhere (perhaps at the pre-processing step?)
 \section*{Title page}
-\textbf{Article title}: The relative velocity of climate change
+\textbf{Article title}: The realised velocity of climate change
 reveals remarkable idiosyncrasy of species' distributional shifts
@@
 
-$^2$ Department of Biological Sciences, Universit\'{e} du Qu\'{e}bec
-\`{a} Montr\'{e}al, Montr\'{e}al, Qu\'{e}bec, Canada
+$^2$ Department of Biological Sciences, University of British
+Columbia, Vanouver, Canada
 
@@
   fungi). We characterise species' responses using a metric that
-  describes the biotic velocity of climate change: how closely
+  describes the realised velocity of climate change: how closely
   species' responses have tracked changing climate through time
-  climate. In contrast to existing metrics that focus purely on space,
-  we focus on species, and thus can examine drivers of interspecific
-  variation. Here we show that species responses are highly
-  idiosyncratic, and thus previous analyses assuming uniform responses
-  may be misleading. This is in stark contrast to species' static
-  present-day and past distributions, which show strong evidence of
-  the imprint of evolutionary history and functional traits. Our
-  analyses are a first step in exploring the vast wealth of empirical
-  data on species' historic responses to recent climate change. We
-  highlight that our findings, while surprising, are consistent with
-  paleontological patterns observed at the onset of mass extinction
-  events.}
-% Is that last sentence a bit much?
+  climate. In contrast to existing `climate velocity' metrics that have focused on space,
+  we focus on species and their null expectation of change
+  % The null expectation bit is new; I think this is a novelty, no?
+  , and thus can examine drivers of inter-specific variation. Here we
+  show that species are tracking climate on average, but not
+  sufficiently to keep up with the pace of climate change. Further,
+  species responses are highly idiosyncratic, and thus previous
+  projections assuming uniform responses may be misleading. This is in
+  stark contrast to species' present-day and historical niche
+  dimensions, which show strong evidence of the imprint of
+  evolutionary history and functional traits. Our analyses are a first
+  step in exploring the vast wealth of empirical data on species'
+  historic responses to recent climate change.}
 
@@
 the Earth---has responded predictably. Present-day extinction rates
-are estimated to be three orders of magnitude greater than background
+are estimated to be up to three orders of magnitude greater than background
 rates \supercite{Pimm2014} and are projected to increase further over
 the next several decades \supercite{MEA2005}. Current estimates
-suggest that over at least one million species may be threatened with
+suggest that over one million species may be threatened with
 extinction\supercite{IPBESland2019}, and that we have experienced a
-50\% decline in animal diversity over the past 40 years
+50\% decline in animal diversity in the past 40 years
 \supercite{McLellan2014}.
@@
 impacts of climate change on biodiversity have attracted much
-attention, and it has been suggested that up to 37\% of species may be
-committed to global extinction by 2050\supercite{Thomas2004}. In the
-face of changing climate species must either shift in space, to track
-favourable conditions, or time, for example, flowering and breeding
-earlier, or migrating sooner. Evidence suggests they are doing both
-\supercite{Parmesan2003,Parmesan2006,Menzel2006,Chen2011}, but not all
-species are responding equally. Some species are expanding their
-distributions, sometimes invasively \supercite{Hellmann2008}, and some
-species are flowering later\supercite{Cook2012}; species may also be
-simply becoming less predictable \supercite{Pearse2017phenology}.
-% I feel this isn't just a token cite; we want to set up our result a
-% little, I think
-Developing robust predictions of ecological responses to climate
-change thus requires data from diverse species, ecosystems, and
-climates to account for these varying drivers. By contrast, current
-projections frequently assume a simplistic or uniform response, and
-have tended to overlook interspecific variation.
-% I'm not sure I necessarily agree with this statement
+attention, but the uncertainty around the magnitude of future
+extinctions (\emph{e.g.}, 37\%\supercite{Thomas2004} vs.\
+7.9\%\supercite{Urban2015} of species) means more work is needed.  In
+the face of changing climate species must either shift in space, to
+track favourable conditions, or time, for example, flowering and
+breeding earlier, or migrating sooner. Evidence suggests they are
+doing both \supercite{Parmesan2003,Parmesan2006,Menzel2006,Chen2011},
+but not all species are responding equally. Some species are expanding
+their distributions, sometimes invasively \supercite{Hellmann2008},
+and some species are flowering later\supercite{Cook2012}; species may
+also be simply becoming less predictable
+\supercite{Pearse2017phenology}, perhaps as evolved responses to
+environmental cues break down.  Simply quantifying species' mean
+responses ignores the important variability in responses among species
+\supercite{Wolkovich2012}. Developing robust predictions of ecological
+responses to climate change thus requires data from diverse species,
+ecosystems, and climates to account for this large inter-specific
+variation.
+% Better, I think, thank you. I think there was something about the
+% previous phrasing that sounded a bit too much like we were knocking
+% people's stats, rather than the concepts, if that makes sense?
 
-We currently lack robust estimates of how species are shifting their
-distributions in response to recent warming. The velocity of
+The velocity of
 temperature change---the rate a species would have to shift its range
@@
 the average pole-ward shift in species' distributions has been
-estimated as approximately 1.69 km/yr-1 \supercite{Chen2011},
+approximately 1.69 km/yr \supercite{Chen2011},
 suggesting that species might be capable of keeping pace with shifting
-climate. Yet this estimate is derived from just a handful of studies,
-and does not address the large variation in response among species.
-Here we leverage the vast volume of distributional data stored in
-digital data repositories\supercite{url_gbif} to provide the first
-global synthesis of species distributional responses to climate
-change. By contrasting historical niche dimensions with those of the
-present day, we can characterise the true biotic velocity of climate
-change, which represents the actual velocity with which species track
-climate across space.
+climate.
+% You questioned the 1.69 figure - this text is from you! :D
+% - I agree, though, that I'm not sure this estimate is fair. I looked
+%   in the supplement and there's a remarkable lack of data going into
+%   the estimate, to be honest with you... Is this sufficiently polite
+%   to address that? They also don't address the file-drawer effect at all.
+Yet this estimate is derived from just 21 studies with taxonomic bias
+reflecting the state of the field at the time (361 birds, 9 mammals,
+and no plants)
+% I want to make sure we're nice about this point, because it's a
+% meta-analysis so it's hardly their fault! Also one of my PhD
+% supervisors is on this paper :p
+, and critically not address the large variation in response among
+species or within taxa. Here we leverage the vast volume of
+distributional data stored in digital data
+repositories\supercite{url_gbif} to provide the first global synthesis
+of species distributional responses to climate change. By contrasting
+historical niche dimensions with those of the present day, we can
+characterise the realised biotic velocity of climate change, which
+represents the actual velocity with which species track climate across
+space.
 
-We show that, on average, species' historical (pre-1970) ranges have
+We show that, on average, species' historical (pre-1980) ranges have
 warmed significantly, and that species have generally shifted their
-distributions to ameliorate this warming. Yet this general pattern
-masks a remarkably degree of variability in species responses: many
-species' ranges are now even warmer than if their distributions had
-remained static. Further, we find that, while species' distributions
-are strongly predicted by their evolutionary history and functional
-traits, the same is not true of their degree of tracking. Thus while
-species' static distributions are reasonably predictable, their
-changes in distributions are not.
+distributions to ameliorate this warming, but not sufficiently to keep
+pace with climate change. Yet this general pattern Yet this general
+pattern masks a remarkably degree of variability in species responses:
+many species' ranges are now even warmer than if their distibutions
+had remained static. Further, we find that, while species' niche
+dimensions are strongly predicted by their evolutionary history and
+functional traits, the same is not true of their degree of
+tracking. Thus while species' distributions are reasonably
+predictable, their changes in distributions are not.
 
@@
 \begin{equation}
-  track_\tau = \frac{present_\tau - past_\tau}{projected_\tau - past_\tau}
+  track_\tau = \frac{current_\tau - past_\tau}{projected_\tau - past_\tau}
 \end{equation}
@@
 represent a change in climate relationship greater than if the species
-had not moved at all. Uncertainty in either species' past or present
+had not moved at all. Uncertainty in either species' past or current
 ranges can be accounted for through a bootstrap procedure, whereby the
-past and/or present distributions are re-sampled (with replacement) to
+past and/or current distributions are re-sampled (with replacement) to
 match the sampling of the other time-period, and the standard
@@
 Our approach has two main innovations: its focus on (1) the whole of a
-species' distribution and(2) expected change. (1) We suggest that
-something of a disconnect is developing among related fields of
-ecology: species' range modellers typically focus on species'
+species' distribution and (2) expected change. (1) Species' range modelers typically focus on species'
 limits\supercite{Sexton2009} (\emph{i.e.}, the upper and lower
@@
 macro-evolutionary \supercite{Zanne2018} biologists focus on central
-tendencies (\emph{i.e.}, the median). We therefore see no strong
+tendencies (\emph{i.e.}, the median). We see no strong
 \emph{a priori} reason that species can, or must, track all quantiles
-equally, and so our approach is to check all quantiles of a species'
-distribution. (2) The denominator of our index re-scales species'
-change around an expectation under distributional stasis. We suggest
-that species' changes should be measured on the basis of the magnitude
-of change to which they are exposed. This change reflects the
-magnitude of ecological and evolutionary selection, and while we
-suggest there are many ways of addressing this, we suggest ours is the
-simplest and so perhaps a fair starting point. We note that our index
-could by re-scaling according to the distance that would need to be
-travelled to match the degree of change (\emph{sensu}
-\citeauthor{Loarie2009}\supercite{Loarie2009}), but prefer to present
-here a dimensionless (and so unit-less) index of tracking.
-% ...something like this? I re-wrote everything to be about the
-% "relative velocity of change" (which would have given us some fancy
-% notation if we wanted it) but as we don't use distances or speeds I
-% feel we ought not to?
+equally. Indeed some evidence suggests different constraints on species
+warm range limts versus their cool range limits, with biotic interactions more influential in the former,
+and climate determining the latter\supercite{Freeman2018}.
+It is possible, therefore, that shifts at species' range edges may be disconnected
+from changes towards the centre of their distribution.
+% phrase better?
+Our approach is to check all quantiles of a species' distribution. (2)
+The denominator of our index re-scales species' change around an
+expectation under distributional stasis. We suggest that species'
+distributional responses should be measured on the basis of the
+magnitude of change to which they are exposed. This change reflects
+the magnitude of ecological and evolutionary selection, and while
+there are many ways of addressing this, we suggest ours is perhaps the
+simplest and most transparent, so perhaps represents a fair starting
+point. We note that we could also rescale our index according to the
+distance that would need to be travelled to maintain climate stasis
+(\emph{sensu} \citeauthor{Loarie2009}\supercite{Loarie2009}), but
+prefer to present here a dimensionless (and so unitless) index of
+tracking, which is simpler to compare across taxa and between
+biogeographic regions.
+% You ask if we could make a map of this. We could, but I don't think
+% we could by Tuesday...
 
@@
 species' distributions (GBIF\supercite{url_gbif}) and climate (UEA's
-CRU\supercite{Harris2014}). We focus on five quantiles: the \nth{5},
-\nth{25}, \nth{50} (median), \nth{75}, and \nth{95}. In the methods we
-give more detail of our reproducible analysis pipeline, and in the
-supplementary materials we repeat our analyses with different data
-curation and cleaning choices. Here, however, we present species'
-range data collected from 1955--2015 for 10,700 species (4,879 plants,
+CRU\supercite{Harris2014}). This represents over 320 Gb of occurrence
+data, collected over the past century.
+% Perhaps now you see why I keep having panic attacks about running
+% everything through!...
+We focus on five quantiles: the \nth{5},
+\nth{25}, \nth{50} (median), \nth{75}, and \nth{95}. Here, we present species' range data collected from 1955--2015 for 10,700 species (4,879 plants,
 1,517 fungi, 273 mammals, 433 birds, 147 reptiles and amphibians, and
@@
 variables in the year in which they were sampled (mean, minimum,
-maximum temperature, diurnal temperature range, precipitation, wet-day
-frequency, vapour pressure, and cloud cover). Our metric was designed
-to account for known biases in the GBIF data \supercite{Beck2014}, and
+maximum temperature, diurnal temperature range, precipitation, 'rainy-day'
+frequency, vapour pressure, and cloud cover). In the methods we
+give more detail of our reproducible analysis pipeline, and in the
+supplementary materials we repeat our analyses with different data
+curation and cleaning choices. Our metric was designed to account for known biases in the GBIF data \supercite{Beck2014}, and
 we present in the supplementary materials additional analyses that
@@
 
-Figure \ref{violin} shows the empirical index values for our focal
-taxa for mean temperature. Clear patterns emerge: 99.5\% of species
-are tracking some climate variable through time, and a large number of
+Figure \ref{violin} shows observed index values for mean temperature.
+Almost all species, 99.5\%, are tracking some climate variable through time, and a large number of
 species are, to some extent, tracking temperature through time (65\%
@@
 movement and a moderate degree of subsequent climate filtering.
+% Jonathan is getting confused about the numbers
+% 99% vs. XXX%; be clearer
 
@@
 29 combinations. This confirms the known results that species'
-present-day and past ranges are strongly influenced by evolutionary,
+current-day and past ranges are strongly influenced by evolutionary,
 but suggests that the magnitude of change (relative to expectations)
@@
 indices are also independent. As we show in figure \ref{pca}, a
-principal components analysis of species' present, past, and projected
+principal components analysis of species' current, past, and projected
 climate distributions shows strong correlations across climate
@@
 species' climate relationships are changing, and thus the factors that
-once are strong predictors of past and present distributions are not
+once are strong predictors of past and current distributions are not
 predictive of the differences between them. A likely
@@
 \supercite{xxx}. (3) If species were perfectly tracking climate
-(\emph{i.e.}, $present_\tau$ is equal to $past_\tau$ in the absence of
+(\emph{i.e.}, $current_\tau$ is equal to $past_\tau$ in the absence of
 sampling error or some other stochastic process) then our index values
@@
 
-This consensus of species' past and present ranges broadly tracking
+This consensus of species' past and current ranges broadly tracking
 the same underlying climate optima is complicated by paleo-ecological
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{../figures/pca-50.pdf}
+    \begin{subfigure}{.45\textwidth}
+      \includegraphics[width=\textwidth]{../figures/pca-05-past.pdf}
+      \caption{Past climate space (1955--1980)}
+    \end{subfigure}
+    \begin{subfigure}{.45\textwidth}
+      \includegraphics[width=\textwidth]{../figures/pca-05-present.pdf}
+      \caption{Present climate space (1990--2015)}
+    \end{subfigure}
+    \begin{subfigure}{.45\textwidth}
+      \includegraphics[width=\textwidth]{../figures/pca-05-index.pdf}
+      \caption{Tracking climate space}
+    \end{subfigure}
   \end{center}
+  % Note: bad filenames I know, but these are the 50th (i.e., median) entries
   \caption{\doublespacing \textbf{Species' movement through climate
@@
     our figures is a principal component analysis (PCA) biplots; in
-    (a) of species' median past climate values, in (b) median present
+    (a) of species' median past climate values, in (b) median current
     climate, and in (c) the median track-index of all climate
@@
 
-All analyses were conducted in \texttt{R} version \texttt{3.6.1}
-\autocite{R2018}. Code to reproduce all analyses in their entirety is
+All analyses were conducted in \emph{R} version \texttt{3.6.1}
+\autocite{R2018}, and all names in \emph{italics} refer to \emph{R}
+packages. Code to reproduce all analyses in their entirety is
 available in the online supplementary materials.
@@
 
-\textbf{Occurrence data.} Species' distribution data were downloaded
-from the Global Biodiversity Information Facility (GBIF)
+Species' distribution data were downloaded from the Global
+Biodiversity Information Facility (GBIF)
 \supercite{url_gbif,url_gbif_occurrences}. We downloaded only data
@@
 contained no obvious location errors (\emph{e.g.}, latitude/longitude
-swaps, coordinates that are rounded and so imprecise). These occurrence
-data include records that were flagged as having potential issues and
-then were corrected; in the supplementary materials, we show that
-excluding these records leads to qualitatively identical results. We
-present, in the main text, results from `human observation'
-occurrences (\emph{e.g.}, human surveys); in the supplementary
-materials observations based upon vouchered specimens (\emph{e.g.},
-pressed specimens in a herbarium) whose results are qualitatively
-identical. We worked only with occurrences from 1955--1980 and
-1990-2015 (see also `metric calculation' below). We only worked with
-species that had at least 1000 records on GBIF. This pragmatic
-threshold limits us to better-studied species, but in preliminary
-investigations we found our results to be robust to this threshold.
+swaps or rounded coordinates). These occurrence data include records
+that were flagged as having potential issues and corrected by GBIF;
+our results were qualitatively identical if we excluded these records
+and so we include them here for completeness. We used `human
+observation' occurrences (\emph{e.g.}, records from surveys), and not
+vouchered specimens (\emph{e.g.}, pressed specimens in a herbarium)
+because preliminary analysis revealed qualitatively identical results
+across the two methods, but with many fewer observations vouchered
+observations. We only worked with species that had at least 1000
+records on GBIF within our focal time periods (1955--1980 and
+1985--2015). This pragmatic threshold limits us to better-studied
+species, but our results are robust to increases (\emph{e.g.}, 10,000
+observations) in this threshold. We only made use of data for which
+species identity was known, and we ignored variation among
+sub-species. The GBIF taxonomy is, itself, to some extent cleaned, and
+so we retained in our analysis any species within GBIF that was not
+present in another datasets (see
+\citeauthor{Cornwell2019}\supercite{Cornwell2019} for a discussion of
+the GBIF taxonomy for plants).
 
-\textbf{Climate data.} Climate data were downloaded from
+Climate data were downloaded from
 \citeauthor{Harris2014}\supercite{Harris2014}'s nine global
-time-series. We split the data into yearly (averaged) $0.5^\circ$
+time-series. We split the data into yearly (averaged) 0.5$^\circ$
 cells, from which we estimated each of our nine climate variables for
-each species observation.
-
-\textbf{Trait and phylogenetic data.} Plant trait data were taken from
+each species observation using
+\emph{raster}\supercite{Hijmans2019}. Plant trait data were taken from
 \citeauthor{Wright2004}\supercite{Wright2004} (leaf lifespan, leaf
@@
 mammal\supercite{Faurby2015}, bird\supercite{Jetz2012}, amphibian
-\supercite{Jetz2018}, and plant\supercite{Smith2018} phylogenies,
-which contain posterior distributions of possible phylogenies which we
-used to account for uncertainty (see `Trait and phylogenetic analyses'
-below). For reptiles, we used
-\citeauthor{Zheng2016}\supercite{Zheng2016}, which is a single
-phylogeny and so we could not account for uncertainty with it in this
-way. We were unable to find dated, global phylogenies for the other
-taxa in this study.
-
-\textbf{Taxonomic cleaning.}  We only made use of data for which
-species identity was known, and we ignored variation among
-sub-species. Species' taxonomy has changed through time, and so is a
-potential source of bias in our analysis. As such, we used the World
-Conservation Union's species taxonomy, through
-\emph{taxize}\supercite{Chamberlain2013}, to harmonise all taxonomic
-information. In cases where a species' name could not be found on the
-IUCN, we retained that species in our analysis (since GBIF itself
-performs taxonomic cleaning and standardisation).
+\supercite{Jetz2018}, plant (`ALLOTB' phylogeny
+\supercite{Smith2018}), and reptiles\supercite{Zheng2016} phylogenies,
+and in cases where posterior distributions of trees were available we
+used a single draw from that distribution.
 
@@
 
-\textbf{Bootstrapping procedure to account for sampling bias and
-  empirical calculation of track index values.}  The sampling of
-species through time and across space is known to be uneven across the
-data in GBIF \supercite{Beck2013,Beck2014}. To ensure our method was
-statistically robust to such changes, we employed a bootstrap
-procedure during the calculation of $track_\tau$. To do this we
-randomly re-sampled, with replacement, the occurrences making up
-$present_\tau$ and $projected_\tau$ to be of the same number as
-$past_\tau$, and vice-versa for $past_\tau$ and $present_\tau$. We
-repeated this process 999 times, and report the median of the
-bootstrapped distribution of $track_\tau$ here in the main text. We
-applied this procedure to the occurrence data described above.
+The sampling of species through time and across space is known to be
+uneven across the data in GBIF \supercite{Beck2013,Beck2014}. To
+ensure our method was statistically robust to such changes, we
+employed a bootstrap procedure during the calculation of
+$track_\tau$. To do this we randomly re-sampled, with replacement, the
+occurrences making up $current_\tau$ and $projected_\tau$ to be of the
+same number as $past_\tau$, and vice-versa for $past_\tau$ and
+$current_\tau$. We repeated this process 999 times, and calculated a
+$track_\tau$ value for each re-sample, generating 999 $track_\tau$
+values whose medians we report here.
 
-\textbf{Estimating the track index in simulated data.} To assess the
-performance of our (bootstrapped) track index, we simulated species'
-whose ranges tracked climate to varying extents through time. Full
-results of the results of these simulations are given in the
-supplementary materials, and a selection of simulations are shown i
-figure \ref{violin}. Using the climate data from above in the years
-1962 and 2002 (the mid-points of the ranges of our data), we simulated
-species with varying maximum possible range sizes (1x1, 2x2, 5x5,
-10x10, 20x20, or 50x50 grid-cell extents) and varying possible
-latitudinal shifts in range ($\pm$1, 2, 5, or 10 grid-cells). We also
+To assess the performance of our (bootstrapped) track index, we
+simulated species' whose ranges tracked climate to varying extents
+through time. Using the climate data from above in the years 1962 and
+2002 (the mid-points of the ranges of our data), we simulated species
+with varying maximum possible range sizes (1x1, 2x2, 5x5, 10x10,
+20x20, or 50x50 grid-cell extents) and varying possible latitudinal
+shifts in range through time ($\pm$1, 2, 5, or 10 grid-cells). We also
 varied two other parameters: a species' degree of environmental
-tracking ($\alpha$; taking a value of 0, 0.1, 0.2, ..., 0.9, or 1) and
-overall fraction of chance occupancy ($\sigma$; taking a value of 0,
-0.1, 0.2, ..., 0.9, or 1). Together, these two parameters define a
-species' probability of being present within a cell within its
-potential range:
+tracking ($\alpha$; taking a value of 0, 0.5, or 1) and overall
+fraction of chance occupancy ($\sigma$; taking a value of 0.5, .75, or
+1). Together, these two parameters define a species' probability of
+being present within a cell within its potential range:
 $(1-\alpha) \times sigma + \alpha \times \mathcal{N}$.  $\mathcal{N}$
@@
 We calculated the Pearson's correlation between all estimated climatic
-niche variables ($present_\tau$, $past_\tau$, $projected_\tau$,
+niche variables ($current_\tau$, $past_\tau$, $projected_\tau$,
 $track_\tau$ and the bootstrapped $track_\tau$). While we show only a

### 2019-10-27 src/analysis-pca.R

@@ -3,46 +3,51 @@ source("src/headers.R")
 
 # Wrapper functions
 .combined.plot <- function(quantile, dims=1:2, clean=5, abs=FALSE){
-    .pca.plot <- function(pca, dims, scale=5, ...){
-        density <- kde(pca$x[,dims])
-        plot(pca$x[,dims], type="n", axes=FALSE)
-        axis(1)
-        axis(2)
-        plot(density, display="filled.contour", cont=seq(10,90,10), cols=heat.colors(length(seq(10,90,10))), add=TRUE, axes=FALSE)
-        points(pca$x, pch=20, cex=.5, col="grey60")
-        arrows(rep(0,ncol(pca$x)),rep(0,ncol(pca$x)), pca$rotation[,dims[1]]*scale, pca$rotation[,dims[2]]*scale, length=.1, lwd=3)
-        #arrows(rep(0,ncol(pca$x)),rep(0,ncol(pca$x)), -pca$rotation[,dims[1]]*scale, -pca$rotation[,dims[2]]*scale, length=.1, lwd=3)
-        text.locs <- pca$rotation*scale*1.3
-        text(text.locs[,dims[1]], text.locs[,dims[2]], names(pca$center), font=2)
-    }
 
-    index <- .simplify.group("bootstrap.index", quantile, clean=clean,  abs=abs)[,1:9]
-    present <- .simplify.group("present", quantile, abs=abs)[,1:9]
-    past <- .simplify.group("past", quantile, abs=abs)[,1:9]
-    browser()
+    index <- na.omit(.simplify.group("bootstrap.index", quantile, clean=clean,  abs=abs)[,1:9])
+    present <- na.omit(.simplify.group("present", quantile, abs=abs)[,1:9])
+    past <- na.omit(.simplify.group("past", quantile, abs=abs)[,1:9])
     
     #layout(matrix(1:6, nrow=2, byrow=TRUE), widths=c(5,5,5), heights=c(5,1))
-    par(mar=c(1,1,1,1), mfrow=c(1,3))
+    par(mar=c(3.1,3.1,1.1,1.1), mfrow=c(1,3))
     present <- prcomp(present, scale=TRUE)
     past <- prcomp(past, scale=TRUE)
     index <- prcomp(index, scale=TRUE)
     .pca.plot(past, 1:2, bty="o")
     .pca.plot(present, 1:2)
     .pca.plot(index, 1:2)
[…]

### 2019-10-27 src/analysis-traits.R

@@ -133,10 +133,10 @@ pres.p <- c(
 )
 sum(pres.p < .05); length(pres.p)
 past.p <- c(
-    .corr.mat("pastent", "0.05", clean=NA, p.val=TRUE),
-    .corr.mat("pastent", "0.25", clean=NA, p.val=TRUE),
-    .corr.mat("pastent", "0.5", clean=NA, p.val=TRUE),
-    .corr.mat("pastent", "0.75", clean=NA, p.val=TRUE),
-    .corr.mat("pastent", "0.95", clean=NA, p.val=TRUE)
+    .corr.mat("past", "0.05", clean=NA, p.val=TRUE),
+    .corr.mat("past", "0.25", clean=NA, p.val=TRUE),
+    .corr.mat("past", "0.5", clean=NA, p.val=TRUE),
+    .corr.mat("past", "0.75", clean=NA, p.val=TRUE),
+    .corr.mat("past", "0.95", clean=NA, p.val=TRUE)
 )
 sum(past.p < .05); length(past.p)

### 2019-10-27 src/analysis-violin.R

@@ -1,82 +1,45 @@
 # Headers
 source("src/headers.R")
 
-# Load data
+# Load empirical data
 plants <- readRDS("clean-data/plants-index.RDS")
 fungi <- readRDS("clean-data/fungi-index.RDS")
 insects <- readRDS("clean-data/insects-index.RDS")
 mammals <- readRDS("clean-data/mammals-index.RDS")
 reptiles <- readRDS("clean-data/reptiles-index.RDS")
 birds <- readRDS("clean-data/birds-index.RDS")
+amphibians <- readRDS("clean-data/amphibians-index.RDS")
+
+# Load simulated data
+load("src/sim-range.RData")
+mid <- t(sapply(results, function(x) x[1,]))
+data <- cbind(data, mid)
+data <- data[!is.na(data$bootstrap.index),]
+sim.data <- data[is.finite(data$index),]
 
 # Do the plots
-.violin.plot <- function(climate, quant, clean=NA, se.thresh=5, abs=FALSE, index="bootstrap.index"){
-    # Load and match data
-    index <- .simplify.group(index, quantile=quant, clean=clean, abs=abs)
-    se <- .simplify.group("mad.index", quantile=quant, abs=abs)
-    se <- se[se[,climate] < se.thresh,]
-    se <- se[rownames(se) %in% rownames(index),]
-    index <- index[rownames(index) %in% rownames(se),]
-    data <- data.frame(index=index[,climate], se=se[,climate], taxon=index$taxon)
-
-    # Prepare colours and order for plotting
-    cols <- setNames(c("forestgreen", "black", "blue", "orange", "red", "grey60"), unique(index$taxon))
-    data$col <- alpha(cols[data$taxon], alpha=1-((data$se^2)/max(data$se^2)))
-    data <- data[order(data$taxon, data$index),]
-    data$x <- as.numeric(factor(data$taxon))
-    counts <- table(data$taxon)
-    curr <- 0
-    for(i in seq_along(counts)){
-        taxon <- names(counts)[i]
-        data$x[data$taxon==taxon] <- data$x[data$taxon==taxon] + seq(-.3, .3, length.out=counts[i])
-    }
-
-    # Lookups for nicer names
[…]

### 2019-10-27 src/download-gbif.rb

@@ -4,11 +4,12 @@
 # data elsewhere
 
 Dir.chdir "~/bear-share/gbif/"
-def download_wrapper(gbif_id, taxon, clean_csv=true, clean_zip=true)
+def download_wrapper(gbif_id, taxon, clean_csv=true, clean_zip=true, clean_gbif=false)
   Dir.chdir taxon
   `wget http://api.gbif.org/v1/occurrence/download/request/#{gbif_id}.zip`
   `unzip #{gbif_id}.zip #{gbif_id}.csv`
   `cut -f13,17,18,28 #{gbif_id}.csv > #{gbif_id}_abbreviated.txt`
+  `cut -f13,17,18,28,44 #{gbif_id}.csv > #{gbif_id}_short_issues.txt`
   if clean_csv then `rm #{gbif_id}.csv` end
   if clean_zip then `rm #{gbif_id}.zip` end
   Dir.chdir '..'
@@ -24,5 +25,4 @@ download_wrapper "0034726-180508205500799" "insecta"
 download_wrapper "0000171-180412121330197" "mammalia"
 download_wrapper "0001849-180131172636756" "plantae"
 download_wrapper "0001851-180131172636756" "plantae"
-download_wrapper "0001847-180131172636756" "pompilidae"
 download_wrapper "0034728-180508205500799" "reptillia"

### 2019-10-27 src/headers.R

@@ -35,6 +35,10 @@ if(!require("viridis")){
     install.packages("viridis"); library(viridis)}
 if(!require("plotrix")){
     install.packages("plotrix"); library(plotrix)}
+if(!require("png")){
+    install.packages("png"); library(png)}
+if(!require("grid")){
+    install.packages("grid"); library(grid)}
 
 ####################################
 # Global settings ##################
@@ -66,8 +70,10 @@ prog.bar <- function(x, y){
                 tryCatch(if(z[1] < 1) if((length(z) %% 10)==0) cat("|") else cat("."), error=function(z) cat("."))
         }
 }    
-.load.gbif <- function(obs, specimens, min.records=2000, clean.binomial=FALSE){
+.load.gbif <- function(obs, specimens, min.records=2000, clean.binomial=FALSE, clean.gbif=FALSE){
     data <- data.table(scientificname="", decimallatitude=0.0, decimallongitude=0.0, year=0, type="")
+    if(clean.gbif)
+        data$issue <- ""
     if(!missing(obs)){
         tmp <- fread(obs)
         tmp$type <- "observations"
@@ -82,6 +88,9 @@ prog.bar <- function(x, y){
         stop("Must supply *some* data to load!...")
     data <- data[-1,]
 
+    if(clean.gbif)
+        data <- data[issue == "",]
+    
     if(clean.binomial){
         spp <- unique(data$scientificname)
         spp <- setNames(tolower(sapply(strsplit(spp, " "), function(x) paste(x[1:2], collapse="_"))), spp)
@@ -204,6 +213,7 @@ prog.bar <- function(x, y){
         data.frame(.simplify(insects, index, quantile, clean, abs), taxon="insects"),
         data.frame(.simplify(mammals, index, quantile, clean, abs), taxon="mammals"),
         data.frame(.simplify(reptiles, index, quantile, clean, abs), taxon="reptiles"),
+        data.frame(.simplify(amphibians, index, quantile, clean, abs), taxon="amphibians"),
         data.frame(.simplify(birds, index, quantile, clean, abs), taxon="birds")
     ))
 .simplify.quantile <- function(index, var, clean=NA, abs=FALSE)

### 2019-10-29 ms/ms.tex
Ready for bioRxiv

@@
 % To-do before bioRxiv:
-% * p-values in lambda
-% * change years in simulation
-% * check taxa numbers in violin plot
-% * add copyright information for phylopics
-% * add sigma into the range simulation code
-% * not a track index. But what is it?...
-% * can you get more birds into the manuscript? They seem to be being dropped somewhere (perhaps at the pre-processing step?)
+% ! p-values in lambda
+
 \section*{Title page}
@@
 $^*$ To whom correspondence should be addressed:
-\url{will.pearse@usu.edu} and \url{j.davies@ubc.ca}
+\emph{will.pearse@usu.edu} and \emph{j.davies@ubc.ca}
 
-\textbf{Word-count}: XXX (excluding figure legends)
+\textbf{Word-count}: 2634 (excluding figure legends)
 
@@
   response to recent climate warming has been based on a few studies
-  with limited number of species. Here we present a comprehensive,
+  with a limited number of species. Here we present a comprehensive,
   global overview of species' distributional responses to changing
@@
   describes the realised velocity of climate change: how closely
-  species' responses have tracked changing climate through time
-  climate. In contrast to existing `climate velocity' metrics that have focused on space,
-  we focus on species and their null expectation of change
-  % The null expectation bit is new; I think this is a novelty, no?
-  , and thus can examine drivers of inter-specific variation. Here we
-  show that species are tracking climate on average, but not
-  sufficiently to keep up with the pace of climate change. Further,
-  species responses are highly idiosyncratic, and thus previous
-  projections assuming uniform responses may be misleading. This is in
-  stark contrast to species' present-day and historical niche
-  dimensions, which show strong evidence of the imprint of
-  evolutionary history and functional traits. Our analyses are a first
-  step in exploring the vast wealth of empirical data on species'
-  historic responses to recent climate change.}
+  species' responses have tracked changing climate through time. In
+  contrast to existing `climate velocity' metrics that have focused on
+  space, we focus on species and index their responses to a null
+  expectation of change in order to examine drivers of inter-specific
+  variation. Here we show that species are tracking climate on
+  average, but not sufficiently to keep up with the pace of climate
+  change. Further, species responses are highly idiosyncratic, and
+  thus highlight that projections assuming uniform responses may be
+  misleading.  This is in stark contrast to species' present-day and
+  historical climate niches, which show strong evidence of the imprint
+  of evolutionary history and functional traits. Our analyses are a
+  first step in exploring the vast wealth of empirical data on
+  species' historic responses to recent climate change.}
 
@@
 the Earth---has responded predictably. Present-day extinction rates
-are estimated to be up to three orders of magnitude greater than background
-rates \supercite{Pimm2014} and are projected to increase further over
-the next several decades \supercite{MEA2005}. Current estimates
-suggest that over one million species may be threatened with
+are estimated to be up to three orders of magnitude greater than
+background rates \supercite{Pimm2014} and are projected to increase
+further over the next several decades \supercite{MEA2005}. Current
+estimates suggest that over one million species may be threatened with
 extinction\supercite{IPBESland2019}, and that we have experienced a
@@
 attention, but the uncertainty around the magnitude of future
-extinctions (\emph{e.g.}, 37\%\supercite{Thomas2004} vs.\
-7.9\%\supercite{Urban2015} of species) means more work is needed.  In
-the face of changing climate species must either shift in space, to
-track favourable conditions, or time, for example, flowering and
-breeding earlier, or migrating sooner. Evidence suggests they are
-doing both \supercite{Parmesan2003,Parmesan2006,Menzel2006,Chen2011},
-but not all species are responding equally. Some species are expanding
-their distributions, sometimes invasively \supercite{Hellmann2008},
-and some species are flowering later\supercite{Cook2012}; species may
-also be simply becoming less predictable
-\supercite{Pearse2017phenology}, perhaps as evolved responses to
-environmental cues break down.  Simply quantifying species' mean
-responses ignores the important variability in responses among species
-\supercite{Wolkovich2012}. Developing robust predictions of ecological
-responses to climate change thus requires data from diverse species,
-ecosystems, and climates to account for this large inter-specific
-variation.
-% Better, I think, thank you. I think there was something about the
-% previous phrasing that sounded a bit too much like we were knocking
-% people's stats, rather than the concepts, if that makes sense?
+extinctions (\emph{e.g.}, 7.9\%\supercite{Urban2015} vs.\
+37\%\supercite{Thomas2004} of species) highlights the inadequacy of
+current forecasts.  In the face of changing climate species must
+either shift in space, to track favourable conditions, or time, for
+example, flowering and breeding earlier.
+% I've dropped the migration reference here because that's shifting in
+% both space and time, and so sort of ruins the either/or structure
+Evidence suggests they are doing both
+\supercite{Parmesan2003,Parmesan2006,Menzel2006,Chen2011}, but not all
+species are responding equally. Some species are expanding their
+distributions, sometimes invasively \supercite{Hellmann2008}, and some
+species are flowering later\supercite{Cook2012}; species may also be
+simply becoming less predictable \supercite{Pearse2017phenology},
+perhaps as evolved responses to environmental cues break down.  Simply
+quantifying species' mean responses ignores the important variability
+in responses among species \supercite{Wolkovich2012}. Developing
+robust predictions of ecological responses to climate change thus
+requires data from diverse species, ecosystems, and climates to
+account for this large inter-specific variation.
 
-The velocity of
-temperature change---the rate a species would have to shift its range
-to maintain constant temperature---has been calculated at 0.42 km/yr
-\supercite{Loarie2009}. A meta-analysis of taxonomic groups estimated
-the average pole-ward shift in species' distributions has been
-approximately 1.69 km/yr \supercite{Chen2011},
+The velocity of temperature change---the rate a species would have to
+shift its range to maintain constant temperature---has been calculated
+at an average of 0.42 km/yr \supercite{Loarie2009}. A meta-analysis of
+taxonomic groups estimated the average pole-ward shift in species'
+distributions has been approximately 1.69 km/yr \supercite{Chen2011},
 suggesting that species might be capable of keeping pace with shifting
-climate.
-% You questioned the 1.69 figure - this text is from you! :D
-% - I agree, though, that I'm not sure this estimate is fair. I looked
-%   in the supplement and there's a remarkable lack of data going into
-%   the estimate, to be honest with you... Is this sufficiently polite
-%   to address that? They also don't address the file-drawer effect at all.
-Yet this estimate is derived from just 21 studies with taxonomic bias
-reflecting the state of the field at the time (361 birds, 9 mammals,
-and no plants)
-% I want to make sure we're nice about this point, because it's a
-% meta-analysis so it's hardly their fault! Also one of my PhD
-% supervisors is on this paper :p
-, and critically not address the large variation in response among
-species or within taxa. Here we leverage the vast volume of
-distributional data stored in digital data
-repositories\supercite{url_gbif} to provide the first global synthesis
-of species distributional responses to climate change. By contrasting
-historical niche dimensions with those of the present day, we can
-characterise the realised biotic velocity of climate change, which
-represents the actual velocity with which species track climate across
-space.
+climate. Yet even this estimate is derived from just 21 studies with
+limited taxonomic coverage (361 birds, 9 mammals, and no plants), and
+critically any single point estimate does not reflect the large
+variation in responses among species or within taxa. Here we leverage
+the vast volume of distributional data stored in digital data
+repositories\supercite{url_gbif}, representing over 10,000 species of
+plants, fungi, mammals, birds, reptiles, amphibians, and insects, to
+provide the first global synthesis of species distributional responses
+to climate change. By contrasting historical niche dimensions with
+those of the present day, we can characterise the realised biotic
+velocity of climate change, which represents the relative degree to
+which species track climate across space.
 
@@
 distributions to ameliorate this warming, but not sufficiently to keep
-pace with climate change. Yet this general pattern Yet this general
-pattern masks a remarkably degree of variability in species responses:
-many species' ranges are now even warmer than if their distibutions
-had remained static. Further, we find that, while species' niche
-dimensions are strongly predicted by their evolutionary history and
-functional traits, the same is not true of their degree of
-tracking. Thus while species' distributions are reasonably
-predictable, their changes in distributions are not.
+pace with climate change. Yet this general pattern masks a remarkably
+degree of variability in species responses, and many species' ranges
+are now even warmer than if their distributions had remained
+static. Further, we find that, while species' niche dimensions are
+strongly predicted by their evolutionary history and functional
+traits, the same is not true of their degree of tracking. Thus while
+species' distributions are reasonably predictable, their changes in
+distributions are not.
 
@@
 \subsection*{Results and discussion}
-\textbf{A comparative index of species' niche tracking.} We propose to
-measure the extent to which a species' distribution is tracking
-climate change by standardising its observed change by the change it
-would have experienced had it not shifted its distribution. Formally,
-we define this as:
+\textbf{A comparative index of species' niche tracking.} We propose a
+measure that captures the extent to which a species' distribution is
+tracking climate by standardising its observed niche change by the
+change it would have experienced had it not shifted its
+distribution. Formally, we define this as:
 
@@
 same quantile of a species' range in the past, and $projected_\tau$
-the same quantile of a species' present-day range under past climatic
+the same quantile of a species' past range under present-day climatic
 conditions. This index should be calculated separately for
@@
 ranges can be accounted for through a bootstrap procedure, whereby the
-past and/or current distributions are re-sampled (with replacement) to
-match the sampling of the other time-period, and the standard
-deviation of the index across those samples reported (see methods).
+past and current distributions are re-sampled to standardise sampling
+across time periods (see methods).
 
 Our approach has two main innovations: its focus on (1) the whole of a
-species' distribution and (2) expected change. (1) Species' range modelers typically focus on species'
-limits\supercite{Sexton2009} (\emph{i.e.}, the upper and lower
-quantiles) while community ecologists \supercite{Tilman2004} and
-macro-evolutionary \supercite{Zanne2018} biologists focus on central
-tendencies (\emph{i.e.}, the median). We see no strong
-\emph{a priori} reason that species can, or must, track all quantiles
-equally. Indeed some evidence suggests different constraints on species
-warm range limts versus their cool range limits, with biotic interactions more influential in the former,
-and climate determining the latter\supercite{Freeman2018}.
-It is possible, therefore, that shifts at species' range edges may be disconnected
-from changes towards the centre of their distribution.
-% phrase better?
-Our approach is to check all quantiles of a species' distribution. (2)
-The denominator of our index re-scales species' change around an
-expectation under distributional stasis. We suggest that species'
+species' distribution and (2) expected change. (1) When estimating
+species' range distributions we typically focus on species'
+limits\supercite{Sexton2009} (\emph{e.g.}, upper and lower quantiles)
+while in community ecological \supercite{Tilman2004} and
+macro-evolutionary \supercite{Zanne2018} studies we tend to focus on
+central tendencies (\emph{e.g.}, the median). However, there is no
+strong \emph{a priori} reason that species can, or must, track all
+quantiles equally. Indeed some evidence suggests different constraints
+on species' warm-range limits versus their cool-range limits, with
+biotic interactions more influential in the former, and climate
+determining the latter\supercite{Freeman2018}.  It is possible,
+therefore, that the drivers of species' range edges and centres may be
+diverging under climate change. Our approach is to check all quantiles
+of a species' distribution. (2) The magnitude of experienced and
+projected climate change shows substantial geographical variation
+\supercite{Loarie2009}, and so species' responses might thus be
+expected to also vary in space. We suggest that species'
 distributional responses should be measured on the basis of the
-magnitude of change to which they are exposed. This change reflects
-the magnitude of ecological and evolutionary selection, and while
-there are many ways of addressing this, we suggest ours is perhaps the
-simplest and most transparent, so perhaps represents a fair starting
-point. We note that we could also rescale our index according to the
-distance that would need to be travelled to maintain climate stasis
-(\emph{sensu} \citeauthor{Loarie2009}\supercite{Loarie2009}), but
-prefer to present here a dimensionless (and so unitless) index of
-tracking, which is simpler to compare across taxa and between
-biogeographic regions.
-% You ask if we could make a map of this. We could, but I don't think
-% we could by Tuesday...
+magnitude of change to which they are exposed, and the denominator of
+our index re-scales species' change around an expectation under
+distributional stasis.  This change reflects the magnitude of
+ecological and evolutionary selection, and while there are many ways
+of addressing this, we suggest ours is perhaps the simplest and most
+transparent, so represents a fair starting point. We note that we
+could also rescale our index according to the distance that would need
+to be travelled to maintain climate stasis (\emph{sensu}
+\citeauthor{Loarie2009}\supercite{Loarie2009}), but prefer to present
+here a unitless index of tracking, which is simpler to compare across
+taxa and between biogeographic regions.
 
@@
 species' distributions (GBIF\supercite{url_gbif}) and climate (UEA's
-CRU\supercite{Harris2014}). This represents over 320 Gb of occurrence
-data, collected over the past century.
-% Perhaps now you see why I keep having panic attacks about running
-% everything through!...
-We focus on five quantiles: the \nth{5},
-\nth{25}, \nth{50} (median), \nth{75}, and \nth{95}. Here, we present species' range data collected from 1955--2015 for 10,700 species (4,879 plants,
-1,517 fungi, 273 mammals, 433 birds, 147 reptiles and amphibians, and
-3,451 insects), each observation matched to one of nine climate
-variables in the year in which they were sampled (mean, minimum,
-maximum temperature, diurnal temperature range, precipitation, 'rainy-day'
-frequency, vapour pressure, and cloud cover). In the methods we
-give more detail of our reproducible analysis pipeline, and in the
-supplementary materials we repeat our analyses with different data
-curation and cleaning choices. Our metric was designed to account for known biases in the GBIF data \supercite{Beck2014}, and
-we present in the supplementary materials additional analyses that
-address and account for such biases.
+CRU\supercite{Harris2014}). This represents over 630 million raw
+occurrence records collected over the past century. We focus on five
+quantiles: the \nth{5}, \nth{25}, \nth{50} (median), \nth{75}, and
+\nth{95}. Here, we present species' range data collected from
+1955--2015 for 10,700 species (4,879 plants, 1,517 fungi, 273 mammals,
+433 birds, 147 reptiles and amphibians, and 3,451 insects), each
+observation matched to one of nine climate variables in the year in
+which they were sampled (mean, minimum, maximum temperature,
+precipitation, `rainy-day' frequency, vapour pressure, potential
+evapotranspiration, `frost day' frequency, and cloud cover). In the
+methods we give more detail of our reproducible analysis pipeline, and
+in the supplementary materials provide instructions for repeating our
+analyses with different data curation and cleaning choices. Our metric
+was designed to account for known biases in the GBIF data
+\supercite{Beck2014}, and in the methods we outline our simulations
+which address the influence of sampling uncertainty.
 
-Figure \ref{violin} shows observed index values for mean temperature.
-Almost all species, 99.5\%, are tracking some climate variable through time, and a large number of
-species are, to some extent, tracking temperature through time (65\%
-track at least one quantile, 43\% track at least two, and 4\% track
-all five). The three temperature indices are the least-tracked of any
+Almost all species are tracking some climate variable through time
+(99.5\% have at least one $track_\tau$ between 0 and 1). Most species
+are, to some extent, tracking temperature (see Figure \ref{overview});
+65\% track at least one quantile and 43\% track at least two.
+% Giving an index in units of distance is a good idea; I think it
+% needs some thought, though, because our index is unitless because we
+% divide the units through; I'm not sure the number we get out of this
+% would be directly interpretable like theirs.
+The three temperature indices are, however, the least-tracked of any
 climate variable, while precipitation is the most-tracked (86\% of
@@
 taxa are not tracking climate through time, and many have experienced
-more change than if their distributions had remained static (99.9\% of
-species have `overshot' in at least three quantiles). As we show in
-figure \ref{violin}, and explore further in the supplementary
-materials, these results are consistent with our null simulations
+more change than if their distributions had remained static. Just as
+most species are tracking at least one quantile of one climate
+variable, 99.9\% of species have `overshoot' (have a $track_\tau$ less
+than 0; for example becoming colder in a warming region) in at least
+one quantile of at least one climate variable. As we show in Figure
+\ref{overview}, these results are consistent with our null simulations
 where species' range change is a combination of essentially random
-movement and a moderate degree of subsequent climate filtering.
-% Jonathan is getting confused about the numbers
-% 99% vs. XXX%; be clearer
+movement and a moderate degree of subsequent climate filtering. Thus
+while our results confirm that, by-and-large, species are tracking
+climate change, the magnitude of this varies across species, taxa, and
+aspects of climate (see Figure \ref{overview}). For example, while
+fungi and insects are, on the whole, broadly tracking the median of
+temperature (median $track_{50\%}$ of 0.46 and 0.64, respectively),
+plants and mammals track temperature (median $track_{50\%}$ of 1.16
+and 1.07, respectively) more poorly than precipitation (median
+$track_{5\%}$ of 0.79 and 0.89, respectively).
 
@@
 examined the phylogenetic signal of species' climate quantiles. We
-focused on mammals\supercite{Faurby2015}, birds\supercite{Jetz2012},
-amphibians \supercite{Jetz2018}, reptiles \supercite{Zheng2016}, and
-plants\supercite{Smith2018}, since we were able to find broadly
-inclusive, dated phylogenies for these taxa. We found statistically
-significant evidence of phylogenetic signal across 158 of 180
-taxon-climate-quantile combinations of species' past ranges (and 153
-in the present; see figure \ref{phylogeny}). This is consistent with
-the known patterns of phylogenetic conservatism outlined above. For
-our index, however, we found evidence of phylogenetic signal in only
-29 combinations. This confirms the known results that species'
-current-day and past ranges are strongly influenced by evolutionary,
-but suggests that the magnitude of change (relative to expectations)
-may be driven by different drivers.
-% Will: this is based on lambda *values* and so you need to update
-% this with p-values...
+focused on mammals, birds, reptiles, and plants, since we were able to
+find broadly inclusive, dated phylogenies for these taxa. We found
+strong evidence of phylogenetic signal in species' past (median
+Pagel's $\lambda$\supercite{Pagel1999} of 0.60) and current (median
+$\lambda$ of 0.57) distributions, but limited signal in our index
+(median $\lambda$ of $<$0.001). Past climate showed stronger
+phylogenetic signal than our index in 158 of all 180 (4 taxa, 5
+quantiles, and 9 climate variables) possible taxon-climate-quantile
+combinations (144 for current climate; see Figures \ref{phylo} and
+\ref{traits-signal}). This confirms the known result that species'
+current-day and past ranges are strongly influenced by
+evolution\supercite{Wiens2010}, but suggests that species' realised
+velocity of change (our tracking index) may be associated with
+different factors.
 
@@
 species' distributions and potential to track changing environment. We
-focused on mammals\supercite{Jones2009a,Wilman2014},
-birds\supercite{Wilman2014}, and plants\supercite{Wright2004}, since
-global, openly-released trait datasets exist for these taxa. Again, we
-found statistically significant associations between traits and past
-distributions of 193 of 240 taxon-climate-quantile combinations (203
-in the present at $\alpha_{5\%}$; see figure \ref{traits}), but not
-for between the traits and our tracking index (9 out of 240l
-4\%). While the traits we used are strongly associated with species'
-climate responses\supercite{Wright2004} (hence their strong predictive
-power of species' distributions, but not tracking index), we emphasise
-that we do not examine traits related to dispersal ability. This is
-due to a paucity of open-access dispersal traits
-\supercite{Gallagher2019}, and we highlight that future work should
-address whether our index is explained by such drivers.
-% Another potential problem here is no trait*env interaction, which I
-% sense is also important...
-% WILL - re-do TEMP.INDEX ~ TEMP! THERMOPHILES!!!
+focused on mammals, birds, and plants, since global, openly-released
+trait datasets exist for these taxa. Again, we found statistically
+significant associations between traits and past distributions of 191
+of 270 (at $\alpha_{5\%}$; 6 traits, 9 climate variables, and 5
+quantiles) taxon-climate-quantile combinations (193 for current
+distributions), but not between the traits and our tracking index (13
+combinations or 5\%; see Figure \ref{traits-signal}). While the traits
+we used are strongly associated with species' climate
+responses\supercite{Wright2004,McCain2014} (hence their strong
+predictive power of species' distributions, but not degree of
+tracking), we emphasise that we could not find sufficient data to
+assess the potential impact of dispersal
+ability\supercite{Schloss2012}. These findings, while surprising, are
+in keeping with emerging evidence that species' traits are a
+relatively poor predictor of change at range edges
+\supercite{Angert2011}.
 
-\textbf{Climate space is not the same as tracking space.} Given our
-index is poorly predicted by traditionally important variables, we now
-consider whether index values are predictive of each other. It is well
-known that global climate\supercite{xxx}, and so species' positions
-across the various axes of climate space\supercite{xxx}, are not
-independent, and so we set out to test whether species' tracking
-indices are also independent. As we show in figure \ref{pca}, a
-principal components analysis of species' current, past, and projected
-climate distributions shows strong correlations across climate
-variables. Yet the index does not: species' relative degree of
-tracking is both of much higher dimension, and shows relatively less
-correlation across climates axes. We argue, therefore, that while
+\textbf{Species' movement through climate space is disconnected from
+  their position within it.} Given our index is poorly predicted by
+traditionally important variables, we now consider whether index
+values are predictive of each other. It is well known that the axes of
+global climate are not independent and that they are not changing
+independently\supercite{Harris2014,IPCC2014}, and so we would expect
+species' relative tracking of climate to show similar patterns. As we
+show in Figure \ref{pca}, principal components analysis of species'
+current and past climate distributions, which we refer to as their
+climate space, shows strong correlations across climate variables. Yet
+our index does not; species' relative degree of tracking is both of
+much higher dimension (the amount of variance explained by each
+principal component axis is similar) and shows relatively less
+correlation across climates axes . We argue, therefore, that while
 species' positions in climate space show strong associations, species'
-relative tracking of their position in climate space does not.
+relative tracking of their position in climate space does not. This
+result explains the poor predictive power of species' traits and
+evolutionary history for our index: the ecological and evolutionary
+rules that determine species' climate relationships are
+changing. Thus, in our analysis, we can explain species' current
+distributions only by the degree to which they resemble past
+distributions.
 
-If true, this result would explain the lack of association between
-species' traits and evolutionary history with their degree of climate
-tracking. The ecological and evolutionary rules that determine
-species' climate relationships are changing, and thus the factors that
-once are strong predictors of past and current distributions are not
-predictive of the differences between them. A likely
+There are alternative explanations for our results, of which we
+highlight three: (1) we are ignoring important niche axes, (2) we are
+ignoring biotic interactions, and (3) our index highlights random
+variation in distribution. (1) Our climate variables are not
+all-encompassing, and we ignore other drivers such as geology and
+habitat-type. Yet we do detect strong pattern in current and past
+distributions, suggesting that the absence of pattern in our indices
+is meaningful. (2) Competition and facilitation among species are
+likely strong drivers of species' ability to track
+climate\supercite{Davis1999}. Yet, again, they are surely also likely
+to have driven species' past distributions, and so it is unclear to us
+why only our metric, and not also current and past distributions, is
+so poorly predicted by our trait and phylogenetic data. (3) If species
+were perfectly tracking climate, $current_\tau$ would be equal to
+$past_\tau$ and so our index values could reflect sampling error or
+some other purely stochastic input. Yet such noise, if truly random,
+would mean our empirical index values would be centred at 0 or
+resemble our simulated data, neither or which is the case (see Figure
+\ref{overview}).
 
-Yet there are alternative
-explanations, of which we highlight three: (1) that we have missed
-fundamentally important environmental axes, (2) biotic interactions,
-and (3) that our index highlights random variation in
-distribution. (1) We fully acknowledge that we ignore potentially
-relevant environmental drivers such as geology (and so soil), habitat,
-and all Anthropogenic factors other than climate. We are not alone in
-this simplification\supercite{xxx}, but we detect strong evolutionary
-pattern in the factors we do examine in the data underlying our
-index. (2) Competition and facilitation among species are likely
-strong drivers of species' ability to track climate\supercite{xxx}. We
-acknowledge it is entirely possible that ignoring these factors is the
-driver of our result and, if so, we urge for an acceleration of the
-inclusion of such information into broad-scale analyses
-\supercite{xxx}. (3) If species were perfectly tracking climate
-(\emph{i.e.}, $current_\tau$ is equal to $past_\tau$ in the absence of
-sampling error or some other stochastic process) then our index values
-would essentially be random numbers. There are good empirical evidence
-that this is not the case \supercite{xxx} but, as we show in the
-supplementary materials, simulations suggest out values are
-incompatible with this result.
-
+\clearpage
 \subsection*{Conclusion}
@@
 responses to climate change being distinct from those that generated
-their ranges in the past. That is not to say that we do not detect the
-imprint of ecological and evolutionary history, or that those earlier
-processes are of predictive power for species distributions, but
-rather that to understand change we must consider new drivers. We do
-not know what those drivers are, although we highlight that biotic
-interactions may play a part. We also highlight that our data,
-although comprehensive, are not taxonomically complete (we address
-only XXX species) and are necessarily spatially ($0.5^\circ$ grid
-cells) and temporally (yearly) coarse given the extent (global and 60
-years) of our study. We are not the first to suggest that, for large
-numbers of species, the extent of climate tracking is idiosyncratic
-\supercite{xxx}. These results are, however, consistent with the
-paleo-ecological record. During and after the last ice-age,
-`no-analogue' communities generated previously-unseen species
-assemblages \supercite{xxx}. Once previous mass extinction events were
-underway, species' extinctions became unpredictable
-\supercite{xxx}. Scientists have been warning for decades that we are
-beginning a sixth mass extinction;
-% Soule 1999 but find something earlier
-perhaps we are seeing the end of its beginning.
-
-This consensus of species' past and current ranges broadly tracking
-the same underlying climate optima is complicated by paleo-ecological
-studies that have revealed greater variability of species' historic
-ranges in previous centuries and millennia than previously thought
-\supercite{Williams2007,Veloz2012,Maguire2015}.
+their distributions in the past. This is in keeping with emerging
+evidence from the paleo-ecological literature that species' historic
+ranges were more variable than previously thought, and seemingly
+influenced by different environmental factors than today
+\supercite{Veloz2012,Maguire2015}. During and after the last ice-age,
+for example, `no-analogue' conditions generated previously-unseen
+species assemblages \supercite{Williams2007}. Equally, our observation
+of idiosyncrasy in species' degree of tracking is consistent with
+species' population crashes (and so extinctions) becoming less
+predictable during previous mass extinction events
+\supercite{Levinton2001,Jablonski2004}. We call upon others to examine
+the additional drivers of our metric of relative change, and to extend
+its definition to include more nuanced definitions of climate
+space. We predict, however, that if the current mass extinction event
+continues, it is likely that patterns of idiosyncrasy among species'
+declines and distributional changes will become more common.
 
@@
   \begin{center}
+    \begin{subfigure}{.7\textwidth}
       \includegraphics[width=\textwidth]{../figures/violin-tmp-0-5.pdf}
+      \caption{$track_{50\%}$(mean-temp)}
+    \end{subfigure}
+    \\
+    \begin{subfigure}{.7\textwidth}
+      \includegraphics[width=\textwidth]{../figures/violin-pre-0-05.pdf}
+      \caption{$track_{5\%}$(precipitation)}
+    \end{subfigure}
   \end{center}
-  \caption{\doublespacing \textbf{Most species are tracking global
-      temperature, but there is pronounced variation among species and
-      clades.} The vertical axis shows the \nth{50} quantile of our
-    tracking index of annual mean temperature (boostrapped to account
-    for variation in sampling; see methods). On the left, our seven
-    major taxonomic groups plotted with the median and \nth{25} and
-    \nth{50} quantiles of their distributions highlighted. At the
-    far-right of the plot, output from our simulations of species that
-    are moving but either imperfectly tracking, or not tracking at
-    all, are shown. In the main text, we highlight that the majority
-    of species are tracking global climate, but there is profound
-    variation among species and clades. Fungi, insects, and reptiles
-    are, broadly, tracking the \nth{50} quantile of mean temperature
-    well, while other taxa (\emph{e.g.}, mammals and plants) are not.}
-  \label{violin}
+  \caption{\textbf{Most species are tracking some aspect of climate,
+      but there is pronounced variation among species and clades.} In
+    (a) we show the \nth{50} quantile (median) of our tracking index
+    of mean annual temperature, and in (b) the \nth{5} quantile of our
+    index of precipitation. On the left of each plot, our seven major
+    taxonomic groups plotted (in green for plants and fungi, and in
+    red for animals). At the far-right of (b), we show the output from
+    our simulations of species that are perfectly tracking,
+    imperfectly tracking, or not tracking climate at all (in blue; see
+    Methods for details). To the right of each distribution of points
+    the median (thickest line) and inter-quartile range (thinner
+    lines) are shown as horizontal lines for each distribution. While
+    the majority of species are tracking global climate to some
+    extent, there is is profound variation among species and
+    clades. Figure \ref{overview}a makes clear the variation among and
+    within taxonomic groups; the distributions of fungi, insects, and
+    reptiles are, broadly, tracking changes in mean temperature well,
+    while other taxa (\emph{e.g.}, mammals and plants) are
+    not. Comparing Figures \ref{overview}a and \ref{overview}b makes
+    clear how different taxonomic groups and species may be tracking
+    different aspects of climate; plants are broadly tracking the
+    lower-limit of precipitation, for instance.}
+  \label{overview}
 \end{figure}
@@
   \end{center}
-  \caption{\doublespacing \textbf{Despite strong phylogenetic signal
-      in species' static distributions, species' changes in
-      distributions are not phylogenetically patterned.} This is a
-    single sample from the posterior distribution of plant phylogenies
-    used for our analysis\supercite{Smith2018}. The two oldest
-    branches have been shortened for clarity of presentation
-    (emphasised with dashed lines). Its internal branches are
+  \caption{\textbf{Despite strong phylogenetic signal in species'
+      static distributions, species' changes in distributions are not
+      phylogenetically patterned.} Here we show the plant tree of
+    life, with the two oldest branches shortened for compactness
+    (emphasised with dashed lines). Internal branches are shaded by
     ancestral state reconstructions, assuming Brownian motion, of
-    species' median mean-temperature in the past. Its tips are
+    species' median mean-temperature in the past. Phylogeny tips are
     coloured according to the track index of the same underlying data
-    (compare with figure \ref{violin}). Both are coloured according to
-    the legend at the bottom of the figure. While this is but a single
-    plot (see also figure \ref{correls} and the supplementary
-    materials), this general pattern of strong phylogenetic signal of
-    climate, but not of degree of tracking, is true of almost all of
-    our underlying data (see main text).}
-  \label{pca}
+    (compare with Figure \ref{overview}). This general pattern of
+    strong phylogenetic signal of climate (shown by the coloured
+    pattern to the internal branches), but not of degree of tracking,
+    is characteristic of almost all of our underlying climate data
+    (see also Figure \ref{traits-signal}).}
+  \label{phylo}
 \end{figure}
@@
       \includegraphics[width=\textwidth]{../figures/traits-50.pdf}
-      \caption{Trait correlations}
+      \caption{Trait correlations (Pearson's $\rho$)}
     \end{subfigure}
@@
       \includegraphics[width=\textwidth]{../figures/physig-5.pdf}
-      \caption{Phylogenetic signal}
+      \caption{Phylogenetic signal (Pagel's $\lambda$)}
     \end{subfigure}
   \end{center}
-  \caption{\doublespacing \textbf{Species' static climate variables
-      show strong trait correlations and phylogenetic signal, but
-      their relative degree of tracking does not.} In (a), we show the
-    correlation between the median of each climate variable (rows)
-    with each of our traits (columns); colours are given at the right
-    of the figure. In each cell, the outer, larger box depicts the
-    correlation between species' present-day climate and the trait,
-    while the smaller, inner box the correlation for species' relative
-    tracking (whose size is proportional to intensity of
-    correlation). The 3 (of 54) track-index correlations that are
-    statistically significant (at $\alpha_{5\%}$) are highlighted with
-    crosses. In (b), we show the phylogenetic signal (Pagel's
-    $\lambda$\supercite{Pagel1999}) of the median of each taxonomic
-    group's (columns) median species' climate variables and relative
-    tracking indices (rows). As with (a), in each cell the outer,
-    larger box depicts the correlation between species' present-day
-    climate and the trait, while the smaller, inner box the
-    correlation for species' relative tracking (whose size is
+  \caption{\textbf{Species' static climate variables show strong trait
+      correlations and phylogenetic signal, but their relative degree
+      of tracking does not.} In (a), we show the correlation
+    (Pearson's $\rho$) between the median of each climate variable
+    (rows) with each of our included traits (columns) using colour
+    (see legend at the right). In each cell, the outer, larger box
+    depicts the correlation between species' present-day climate and
+    the trait, while the smaller, inner box the correlation for
+    species' relative tracking (whose size is proportional to
+    intensity of correlation). The 4 (of 54) track-index correlations
+    that are statistically significant (at $\alpha_{5\%}$) are
+    highlighted with crosses. In (b), we show, for each taxon, the
+    phylogenetic signal (Pagel's $\lambda$\supercite{Pagel1999}) of
+    major climate variables and relative tracking indices (rows) using
+    colour (see legend at the right). As with (a), in each cell the
+    outer, larger box depicts the correlation between species'
+    present-day climate and the trait, while the smaller, inner box
+    the correlation for species' relative tracking (whose size is
     proportional to phylogenetic signal strength). The single
@@
     the underlying data is highlighted with a cross.}
-  \label{traits}
+  \label{traits-signal}
 \end{figure}
+% To-do: find an alternative to the crosses
 
@@
   \end{center}
-  % Note: bad filenames I know, but these are the 50th (i.e., median) entries
-  \caption{\doublespacing \textbf{Species' movement through climate
-      space differs from their current positions within it.} Each of
-    our figures is a principal component analysis (PCA) biplots; in
-    (a) of species' median past climate values, in (b) median current
-    climate, and in (c) the median track-index of all climate
-    variables. A PCA biplot shows correlations among the underlying
-    data, with each axis explaining some amount of variation in the
-    underlying data (labelled on the axes themselves). In (a) and (b),
-    species' climate variables are correlated with one-another, and
-    the space reflects global patterns in climate. In (c), this
-    pattern has broken down: while some associations among climate
-    axes still remain (notably temperature), the general pattern is
-    one of independent shift through climate space.}
+  % Note: bad filenames I know, but these are the 50th entries
+  \caption{\textbf{Species' movement through climate space differs
+      from their current and past positions within it.}  Principal
+    component analysis (PCA) biplots of (a) species' median past
+    climate values, (b) median current climate, and (c) the median
+    track-index of all climate variables. Each PCA axis explains some
+    amount of variation in the underlying data (labelled on the axes
+    themselves). In (a) and (b), species' climate variables are
+    correlated with one-another, and the space reflects global
+    patterns in climate. In (c), this pattern has broken down: while
+    some associations among climate axes still remain (notably
+    temperature), the general pattern is one of uncorrelated shifts
+    through climate space.}
   \label{pca}
@@
 was funded by \emph{Fonds de Recherche Nature et Technologies} grant
-number 168004.
+number 168004. The images in Figure \ref{overview} are taken, with
+gratitude, from \emph{http://phylopic.org/}; all are under the Public
+Domain Dedication 1.0 license and uncredited unless otherwise stated
+below. The reptile image was drawn by Jack Mayer Wood, the plant by
+Jason McNair, the insect by Birgit Lang, the mammal by Scott Hartman
+(used under the Creative Commons Attribution 3.0 Unported license),
+and the amphibian by Nobu Tamura (used under the Creative Commons
+Attribution 3.0 Unported license).
 
@@
 
-\textbf{Author information} Reprints and permissions information is
-available at \url{www.nature.com/reprints}. The authors declare no
-competing financial interests. Correspondence and requests for
-materials should be addressed to \url{will.pearse@usu.edu}.
+\textbf{Author information} The authors declare no competing financial
+interests. Correspondence and requests for materials should be
+addressed to \emph{will.pearse@usu.edu} and \emph{j.davies@ubc.ca}.
 
@@
 \section*{\Large Methods}
-\section*{\Large Methods}
 \renewcommand{\figurename}{Extended Data Figure}
@@
 
-All analyses were conducted in \emph{R} version \texttt{3.6.1}
-\autocite{R2018}, and all names in \emph{italics} refer to \emph{R}
+All analyses were conducted in \emph{R} version \emph{3.6.1}
+\supercite{R2018}, and all names in \emph{italics} refer to \emph{R}
 packages. Code to reproduce all analyses in their entirety is
-available in the online supplementary materials.
+available in the online supplementary materials and online at
+\emph{https://github.com/willpearse/track-index}.  In the
+supplementary materials, we provide extended figures summarising all
+tracking index quantiles and climate variables in the same detail as
+presented in the main text.
 
@@
 across the two methods, but with many fewer observations vouchered
-observations. We only worked with species that had at least 1000
-records on GBIF within our focal time periods (1955--1980 and
-1985--2015). This pragmatic threshold limits us to better-studied
+observations.
+
+We only worked with species that had at least 1000 records on GBIF
+within our focal time periods (1955--1980 and 1985--2015). We focused
+on these time periods because they contain sufficient observations for
+large-scale analysis. Since climate change was underway within the
+1980s \supercite{IPCC2014}, this also allows us to have reasonably
+large (if na\"{i}vely defined) `pre-change' and `post-change'
+periods. Our 1000 observation threshold limits us to better-studied
 species, but our results are robust to increases (\emph{e.g.}, 10,000
@@
 species identity was known, and we ignored variation among
-sub-species. The GBIF taxonomy is, itself, to some extent cleaned, and
-so we retained in our analysis any species within GBIF that was not
-present in another datasets (see
+sub-species. The GBIF taxonomy is, itself, to some extent checked and
+harmonised, and so we retained in our analysis any species within GBIF
+that was not present in another datasets (see
 \citeauthor{Cornwell2019}\supercite{Cornwell2019} for a discussion of
-the GBIF taxonomy for plants).
+the GBIF taxonomy for plants). In the supplementary materials, we give
+instructions for re-running our analysis with differing curation
+thresholds to verify our choices.
 
@@
 mass-per-unit-area, leaf Nitrogen mass, and leaf photosynthetic
-capacity), mammal body mass data from PanTHERIA\supercite{Jones2009a}
-and EltonTraits \supercite{Wilman2014}, and bird body mass data from
+capacity), mammal body mass data from the Amniote Trait
+Database\supercite{Myhrvold2015}, and bird body mass data from
 EltonTraits \supercite{Wilman2014}. We were unable to find sufficient
@@
 and in cases where posterior distributions of trees were available we
-used a single draw from that distribution.
+used a single draw from that distribution. Missing species were added
+into the phylogenies using \emph{congeneric.merge} in
+\emph{pez}\supercite{Pearse2015}.
 
 \subsection*{Calculating and testing our track index}
-
-The sampling of species through time and across space is known to be
-uneven across the data in GBIF \supercite{Beck2013,Beck2014}. To
-ensure our method was statistically robust to such changes, we
-employed a bootstrap procedure during the calculation of
-$track_\tau$. To do this we randomly re-sampled, with replacement, the
-occurrences making up $current_\tau$ and $projected_\tau$ to be of the
-same number as $past_\tau$, and vice-versa for $past_\tau$ and
-$current_\tau$. We repeated this process 999 times, and calculated a
-$track_\tau$ value for each re-sample, generating 999 $track_\tau$
-values whose medians we report here.
+As we describe in the main text, we quantify the realised velocity of
+climate change using a simple index of climate tracking. Our index,
+$track_\tau$, scales the observed magnitude of species' shift in
+climate ($current_\tau - past_\tau$) according to the degree of change
+to which that species was exposed ($projected_\tau - past_\tau$). Our
+metric is comparable across species, and can also be calculated in
+such a way as to control for changes in sampling. The sampling of
+species through time and across space is known to be uneven across the
+data in GBIF \supercite{Beck2013,Beck2014}. To ensure our method was
+statistically robust to such changes, we employed a bootstrap
+procedure during the calculation of $track_\tau$. To do this we
+randomly re-sampled, with replacement, the occurrences making up
+$current_\tau$ and $projected_\tau$ to be of the same number as
+$past_\tau$, and vice-versa for $past_\tau$ and $current_\tau$. We
+repeated this process 999 times, and calculated a $track_\tau$ value
+for each re-sample, generating 999 $track_\tau$ values whose medians
+we report here. This process accounts for uneven sampling by
+generating a pseudo-posterior distribution of conservatively-estimated
+values. In each bootstrap, the better-sampled time period is
+sub-sampled to match the poorer-sampled period, accounting for
+differences in sampling. Equally, the poorer-sampled period is
+re-sampled in order to increase the variance around the index,
+essentially cross-validating our estimate. When applied to empirical
+data, our bootstrap approach reveals that the certainty differing
+tracking quantiles (and climate variables) is uneven. We therefore
+excluded uncertain (\emph{i.e.}, estimates with high variation across
+pseudo-posteriors) and outliers from figures within the main text, and
+so report sample sizes within Figure \ref{overview} (and its
+counterparts in the supplementary materials).
 
 To assess the performance of our (bootstrapped) track index, we
-simulated species' whose ranges tracked climate to varying extents
+simulated species' with ranges that tracked climate to varying extents
 through time. Using the climate data from above in the years 1962 and
 2002 (the mid-points of the ranges of our data), we simulated species
-with varying maximum possible range sizes (1x1, 2x2, 5x5, 10x10,
-20x20, or 50x50 grid-cell extents) and varying possible latitudinal
-shifts in range through time ($\pm$1, 2, 5, or 10 grid-cells). We also
-varied two other parameters: a species' degree of environmental
-tracking ($\alpha$; taking a value of 0, 0.5, or 1) and overall
-fraction of chance occupancy ($\sigma$; taking a value of 0.5, .75, or
-1). Together, these two parameters define a species' probability of
-being present within a cell within its potential range:
+with varying maximum possible range sizes (2x2, 5x5, 10x10, or 20x20
+grid-cell extents) and varying latitudinal shifts in range through
+time (-4, -3, ..., 3, or 4 grid-cells). We also varied species' degree
+of environmental tracking ($\alpha$; taking a value of 0, 0.5, or 1)
+and probability of occupancy($\sigma$; taking a value of 0.5, .75, or
+1). Together, these latter two parameters define a species'
+probability of being present in a particular cell within its range:
 $(1-\alpha) \times sigma + \alpha \times \mathcal{N}$.  $\mathcal{N}$
@@
 variance of 1. Thus a species with an $\alpha$ of 1 is present only in
-cells within its range that resemble its climatic centre, and a
-species with an $\alpha$ of 0 is present in proportion $\sigma$ of the
-cells within its potential range. We simulated all possible
-combinations of parameters 20 times, providing a full range of
-potential range shifts.
-% This needs more work!
+cells within its range that resemble its climatic centre (\emph{i.e.},
+are similar to the median value of the species' climatic
+distribution), and a species with an $\alpha$ of 0 is present in
+proportion $\sigma$ of the cells within its potential range. We
+simulated all possible combinations of parameters across 100 possible
+centre-points of species' ranges, providing a full range of potential
+range shifts and 26,598 total simulation runs. As with our empirical
+data, in simulation runs where, by chance, simulated species were not
+present in the past or current time-period, they were excluded from
+the analysis.
 
-\subsection*{Trait and phylogenetic analyses}
+\subsection*{Trait, phylogenetic, and PCA analyses}
 
@@
 $track_\tau$ and the bootstrapped $track_\tau$). While we show only a
-handful of these correlations in figure \ref{traits}, we provide all
-correlations in the supplementary figures (and note that the
-qualitative results of all correlations are the same). For the
+handful of these correlations in Figure \ref{traits-signal}, we
+provide all correlations in the supplementary figures (and note that
+the qualitative results of all correlations are the same). For the
 phylogenetic signal analyses, we likewise calculated Pagel's $\lambda$
 (using \emph{phytools}\supercite{Revell2012phytools}) for all indices,
-reporting all results in full in the supplementary materials (which
-are again qualitatively identical to the results in the main text).
+reporting all results in full in the supplementary materials (results
+are, again, qualitatively identical to the results in the main
+text). Principal component analyses were performed on all current,
+past, and track indices across all variables for a given quantile
+(\emph{i.e.}, the \nth{5}, \nth{25}, \nth{50}, \nth{75}, and
+\nth{95}). While we report only the \nth{50} quantiles in the main
+text, in the supplementary materials we give results for all
+quantiles, which are qualitatively identical to the results in the
+main text.
 

### 2019-10-29 src/analysis-pca.R

@@ -36,6 +36,7 @@ insects <- readRDS("clean-data/insects-index.RDS")
 mammals <- readRDS("clean-data/mammals-index.RDS")
 reptiles <- readRDS("clean-data/reptiles-index.RDS")
 birds <- readRDS("clean-data/birds-index.RDS")
+amphibians <- readRDS("clean-data/amphibians-index.RDS")
 
 # PCAs
 for(quant in c("0.05","0.25","0.5","0.75","0.95")){

### 2019-10-29 src/analysis-phylo.R

@@ -20,14 +20,19 @@ source("src/headers.R")
         apply(as.matrix(c.data$data), 1, function(x) which.min(abs(x))),
         rownames(c.data$data)
     )
-.phy.sig.wrap <- function(data, tree, index, clean, method="lambda", merge=TRUE){
+.phy.sig.wrap <- function(data, tree, index, clean, method="lambda", merge=TRUE, p.val=FALSE){
     output <- matrix(NA, ncol=9, nrow=5)
     rownames(output) <- dimnames(data)[[2]]
     colnames(output) <- dimnames(data)[[1]]
     
     for(i in seq_len(nrow(output))){
         c.data <- .make.comparative.data(.simplify(data, index, rownames(output)[i]), tree, merge)
-        output[i,] <- as.numeric(.phy.sig(c.data)[1,])
+        if(p.val){
+            output[i,] <- as.numeric(.phy.sig(c.data)[2,])
+        } else {
+            output[i,] <- as.numeric(.phy.sig(c.data)[1,])
+        }
+            
     }
     return(output)
 }
@@ -109,7 +114,8 @@ reptiles <- readRDS("clean-data/reptiles-index.RDS")
 birds <- readRDS("clean-data/birds-index.RDS")
 
 # Calculate signals
-saveRDS(list(
+if(FALSE){
+    saveRDS(list(
     bootstrap.index=.phy.sig.wrap(mammals, read.tree("phylogenies/faurby_mammals.tre"), "bootstrap.index"),
     index=.phy.sig.wrap(mammals, read.tree("phylogenies/faurby_mammals.tre"), "index"),
     present=.phy.sig.wrap(mammals, read.tree("phylogenies/faurby_mammals.tre"), "present"),
@@ -145,6 +151,45 @@ saveRDS(list(
     past=.phy.sig.wrap(plants, read.tree("phylogenies/ALLOTB.tre"), "past"),
     projected=.phy.sig.wrap(plants, read.tree("phylogenies/ALLOTB.tre"), "projected")
 ), "clean-data/plants-signal.RDS")
+}
+
+# Get p-values
+saveRDS(list(
+    bootstrap.index=.phy.sig.wrap(mammals, read.tree("phylogenies/faurby_mammals.tre"), "bootstrap.index"),
+    index=.phy.sig.wrap(mammals, read.tree("phylogenies/faurby_mammals.tre"), "index"),
[…]

### 2019-10-29 src/analysis-traits.R

@@ -14,9 +14,11 @@ source("src/headers.R")
 .combine.data <- function(index, quant, clean=10, abs=FALSE){
     data <- .simplify.group(index, quant, clean, abs)
     data$species <- rownames(data)
-
-    data$bm.neotoma <- neotoma$log.mass[match(data$species, neotoma$binomial)]
-    data$bm.elton <- elton.bm[data$species]
+    
+    data$bm.mamm <- neotoma$log.mass[match(data$species, neotoma$binomial)]
+    data$bm.mamm[data$taxon != "mammals"] <- NA
+    data$bm.bird <- elton.bm[data$species]
+    data$bm.bird[data$taxon != "birds"] <- NA
 
     data$log.ll <- glopnet$log.LL[match(data$species, glopnet$Species)]
     data$log.lma <- glopnet$log.LMA[match(data$species, glopnet$Species)]
@@ -28,19 +30,19 @@ source("src/headers.R")
     data <- .combine.data(index, quant, clean, abs)
 
     c.var <- setNames(
-        c("clouds","frost","evapo-trans.","precipitation","min(temp)","mean(temp)","max(temp)","vapour"),
-                      c("cld","frs","pet","pre","tmn","tmp","tmx","vap")
+        c("clouds","frost","evapo-trans.","precipitation","min(temp)","mean(temp)","max(temp)","vapour","rainy day"),
+        c("cld","frs","pet","pre","tmn","tmp","tmx","vap","wet")
     )
     q.var <- setNames(
         c("5th","25th","50th","75th","95th"),
         c("0.05","0.25","0.5","0.75","0.95")
     )[quant]
     t.var <- setNames(
-        c("body-mass","body-mass","leaf lifespan","leaf mass/area","leaf N","photosynth"),
-        c("bm.neotoma","bm.elton","log.ll","log.lma","log.Nmass","log.Amass")
+        c("mammal mass","bird mass","leaf lifespan","leaf mass/area","leaf N","photosynth"),
+        c("bm.mamm","bm.bird","log.ll","log.lma","log.Nmass","log.Amass")
     )
     
-    mat <- matrix(nrow=6, ncol=8, dimnames=list(t.var, c.var))
+    mat <- matrix(nrow=length(t.var), ncol=length(c.var), dimnames=list(t.var, c.var))
     for(i in seq_along(c.var)){
[…]

### 2019-10-29 src/analysis-violin.R

@@ -18,16 +18,16 @@ data <- data[!is.na(data$bootstrap.index),]
 sim.data <- data[is.finite(data$index),]
 
 # Do the plots
-.outline.plot <- function(climate, quant, clean=NA, se.thresh=5, abs=FALSE, index="bootstrap.index"){
+.outline.plot <- function(climate, quant, clean=NA, se.thresh=5, abs=FALSE, index="bootstrap.index", sims=FALSE){
     # Prep simulated data
     #sim.data <- sim.data[as.character(sim.data$quantile)==quant,]
     warning("Currently only have simulated data for 0.5 quantile and temperature")
     sim.data <- sim.data[order(sim.data$alpha),]
     sim.data$index <- sim.data[,index]
-    sim.data$alpha.cat <- ifelse(sim.data$alpha==0.5, "intermediate", ifelse(sim.data$alpha==0, "static", "perfect tracking"))
-    sim.data$j.a.cat <- factor(sim.data$alpha.cat, levels=c("static","intermediate","perfect tracking"))
+    sim.data$alpha.cat <- ifelse(sim.data$alpha==0.5, "intermediate", ifelse(sim.data$alpha==0, "static", "tracking"))
+    sim.data$j.a.cat <- factor(sim.data$alpha.cat, levels=c("static","intermediate","tracking"))
     sim.data$j.a.cat <- as.numeric(sim.data$j.a.cat)
-    
+
     # Load and match data
     index <- .simplify.group(index, quantile=quant, clean=clean, abs=abs)
     data <- data.frame(index=index[,climate], taxon=index$taxon)
@@ -39,6 +39,7 @@ sim.data <- data[is.finite(data$index),]
     # Check for order of names as otherwise plots will be wrong
     if(!identical(names(cols), as.character(unique(data$taxon))))
         stop("Names of taxa in wrong order; plots will be wrong!")
+    
     
     # Lookups for nicer names
     c.var <- setNames(c("cloud-cover","frost-days","potential-evapotranspiration","precipitation","min-temperature","mean-temperature","max-temperature","vapour-pressure"),
@@ -58,46 +59,67 @@ sim.data <- data[is.finite(data$index),]
 
     
     par(mar=c(6.1,5.1,1.1,1.1))
+    if(sims){
[…]

### 2019-10-29 src/headers.R

@@ -223,5 +223,6 @@ prog.bar <- function(x, y){
         data.frame(.simplify(insects, index, quantile, clean, abs), taxon="insects"),
         data.frame(.simplify(mammals, index, quantile, clean, abs), taxon="mammals"),
         data.frame(.simplify(reptiles, index, quantile, clean, abs), taxon="reptiles"),
+        data.frame(.simplify(amphibians, index, quantile, clean, abs), taxon="amphibians"),
         data.frame(.simplify(birds, index, quantile, clean, abs), taxon="birds")
     ))

### 2019-10-29 src/sim-range.R

@@ -15,36 +15,40 @@ temp <- readRDS("clean-data/cru-tmp.RDS")
     !(any(dist < 0) | any(dist[,1] > nrow(raster)) | any(dist[,2] > ncol(raster)))
 .get.niche <- function(dist, raster)
     median(raster[dist], na.rm=TRUE)
-.prob.occur <- function(dist, raster, alpha, sd, optimum){
+.prob.occur <- function(dist, raster, alpha, sd, optimum, sigma){
     env <- raster[dist]
     niche <- dnorm(env, optimum, sd) / dnorm(optimum, optimum, sd)
-    niche <- (1-alpha) + niche*alpha
+    niche <- ((1-alpha)*sigma) + (niche*alpha)
     present <- dist[runif(length(env)) <= niche,]
     return(present)
 }
-.dist.wrap <- function(x, y, range, raster, alpha, sd, optimum){
+.dist.wrap <- function(x, y, range, raster, alpha, sd, optimum, sigma){
     block <- .make.block.dist(x, y, range, raster)
-    return(.prob.occur(block, raster, alpha, sd, optimum))
+    return(.prob.occur(block, raster, alpha, sd, optimum, sigma))
 }
 .do.calc <- function(i){
-    # Make distributions
-    # THIS IS THE ONE
-    optimum <- .get.niche(.make.block.dist(data$x[i], data$y[i], data$range[i], temp), early)
-    early.dist <- .dist.wrap(data$x[i], data$y[i], data$range[i], early, data$alpha[i], 1, optimum)
-    mid.dist <- .dist.wrap(data$x[i], data$y[i], data$range[i], mid, data$alpha[i], 1, optimum)
-    late.dist <- .dist.wrap(data$x[i], data$y[i], data$range[i], late, data$alpha[i], 1, optimum)
-    # Calculate index
-    mid.index <- .calc.metric(as.numeric(mid[mid.dist]), as.numeric(early[early.dist]), as.numeric(mid[early.dist]), quantile=.5, n.boot=99)
-    late.index <- .calc.metric(as.numeric(late[late.dist]), as.numeric(early[early.dist]), as.numeric(late[early.dist]), quantile=.5, n.boot=99)
+    locker <- FALSE
+    tryCatch({
+        # Make distributions
+        optimum <- .get.niche(.make.block.dist(data$x[i], data$y[i], data$range[i], temp), early)
[…]

### 2020-12-08 ms/ms.tex
Pre-revamp

@@
 \usepackage{pdflscape}
+
+\usepackage{xspace}
+\newcommand{\eg}{\emph{e.g.}\xspace}
 
@@
 estimates suggest that over one million species may be threatened with
-extinction\supercite{IPBESland2019}, and that we have experienced a
-50\% decline in animal diversity in the past 40 years
-\supercite{McLellan2014}.
-
-The main direct human-induced drivers that impact biodiversity now are
-habitat loss \supercite{Newbold2015} and fragmentation
-\supercite{Haddad2015}, but climate change is likely to become a
-dominant driver in the next few
+extinction\supercite{IPBESland2019}, and while the main human-induced
+drivers of ongoing biodiversity loss are habitat loss
+\supercite{Newbold2015} and fragmentation \supercite{Haddad2015},
+climate change is likely to become a dominant driver in the next few
 decades\supercite{Parmesan2003,Thuiller2007,Urban2015}. Projected
 impacts of climate change on biodiversity have attracted much
-attention, but the uncertainty around the magnitude of future
-extinctions (\emph{e.g.}, 7.9\%\supercite{Urban2015} vs.\
-37\%\supercite{Thomas2004} of species) highlights the inadequacy of
-current forecasts.  In the face of changing climate species must
-either shift in space, to track favourable conditions, or time, for
-example, flowering and breeding earlier.
+attention, but the uncertainty around the magnitude (\emph{e.g.},
+7.9\%\supercite{Urban2015} vs.\ 37\%\supercite{Thomas2004} of species)
+and timing \supercite{Trisos2020} of future extinctions highlights the
+inadequacy of current forecasts.  In the face of changing climate
+species must either shift in space, to track favourable conditions, or
+time, for example, flowering and breeding earlier.
 % I've dropped the migration reference here because that's shifting in
@@
 \supercite{Parmesan2003,Parmesan2006,Menzel2006,Chen2011}, but not all
-species are responding equally. Some species are expanding their
+species are responding equally and impacts on species vary across
+environments \supercite{Pinsky2019}. Some species are expanding their
 distributions, sometimes invasively \supercite{Hellmann2008}, and some
 species are flowering later\supercite{Cook2012}; species may also be
-simply becoming less predictable \supercite{Pearse2017phenology},
+becoming less predictable \supercite{Pearse2017phenology},
 perhaps as evolved responses to environmental cues break down.  Simply
@@
 critically any single point estimate does not reflect the large
-variation in responses among species or within taxa. Here we leverage
-the vast volume of distributional data stored in digital data
+variation in responses among species or within taxa. Further, there is
+emerging evidence that average temperature may be less important in
+driving local extinctions than changes in extremes of temperature
+\supercite{Roman-Palacios2020} Here we leverage the vast volume of
+distributional data stored in digital data
 repositories\supercite{url_gbif}, representing over 10,000 species of
@@
 pace with climate change. Yet this general pattern masks a remarkably
-degree of variability in species responses, and many species' ranges
-are now even warmer than if their distributions had remained
-static. Further, we find that, while species' niche dimensions are
-strongly predicted by their evolutionary history and functional
-traits, the same is not true of their degree of tracking. Thus while
-species' distributions are reasonably predictable, their changes in
-distributions are not.
+degree of variability in species responses, and a surprising number of
+species' ranges are now even warmer than if their distributions had
+remained static. Further, we find that, while species' niche
+dimensions are strongly predicted by their evolutionary history and
+functional traits, the same is not true of their degree of
+tracking. Thus while species' distributions are reasonably
+predictable, their changes in distributions are not.
 
@@
 conditions. This index should be calculated separately for
-temperature, pollution, or whatever environmental variable is of
+temperature, precipitation, or whatever environmental variable is of
 interest. Thus a $track$ of 0 would indicate a species that has
@@
 
-Our approach has two main innovations: its focus on (1) the whole of a
-species' distribution and (2) expected change. (1) When estimating
-species' range distributions we typically focus on species'
-limits\supercite{Sexton2009} (\emph{e.g.}, upper and lower quantiles)
-while in community ecological \supercite{Tilman2004} and
-macro-evolutionary \supercite{Zanne2018} studies we tend to focus on
-central tendencies (\emph{e.g.}, the median). However, there is no
-strong \emph{a priori} reason that species can, or must, track all
-quantiles equally. Indeed some evidence suggests different constraints
-on species' warm-range limits versus their cool-range limits, with
-biotic interactions more influential in the former, and climate
-determining the latter\supercite{Freeman2018}.  It is possible,
-therefore, that the drivers of species' range edges and centres may be
-diverging under climate change. Our approach is to check all quantiles
-of a species' distribution. (2) The magnitude of experienced and
-projected climate change shows substantial geographical variation
+Our approach has two main innovations: its (1) focus on the whole of a
+species' distribution and (2) comparison of observed change with a
+null expectation. (1) When estimating species' range distributions we
+typically focus on species' limits\supercite{Sexton2009} (\emph{e.g.},
+upper and lower quantiles) while in community ecological
+\supercite{Tilman2004} and macro-evolutionary \supercite{Zanne2018}
+studies we tend to focus on central tendencies (\emph{e.g.}, the
+median). However, there is no strong \emph{a priori} reason that
+species can, or must, track all quantiles equally. Indeed some
+evidence suggests different constraints on species' warm-range limits
+versus their cool-range limits, with biotic interactions more
+influential in the former, and climate determining the
+latter\supercite{Freeman2018}. It is possible, therefore, that the
+drivers of species' range edges and centres may be diverging under
+climate change. Our approach is to check all quantiles of a species'
+distribution. (2) The magnitude of experienced and projected climate
+change shows substantial geographical variation
 \supercite{Loarie2009}, and so species' responses might thus be
@@
 \nth{95}. Here, we present species' range data collected from
-1955--2015 for 10,700 species (4,879 plants, 1,517 fungi, 273 mammals,
-433 birds, 147 reptiles and amphibians, and 3,451 insects), each
-observation matched to one of nine climate variables in the year in
-which they were sampled (mean, minimum, maximum temperature,
-precipitation, `rainy-day' frequency, vapour pressure, potential
-evapotranspiration, `frost day' frequency, and cloud cover). In the
-methods we give more detail of our reproducible analysis pipeline, and
-in the supplementary materials provide instructions for repeating our
-analyses with different data curation and cleaning choices. Our metric
-was designed to account for known biases in the GBIF data
-\supercite{Beck2014}, and in the methods we outline our simulations
-which address the influence of sampling uncertainty.
+1955--2015 for XXX species (XXX plants, XXX fungi, XXX mammals, XXX
+birds, XXX reptiles and amphibians, and XX insects), each observation
+is matched to one of nine climate variables in the year in which they
+were sampled. In the methods we give more detail of our reproducible
+analysis pipeline, and in the supplementary materials provide
+instructions for repeating our analyses with different data curation
+and cleaning choices. Our metric was designed to account for known
+biases in the GBIF data \supercite{Beck2014}, and in the methods we
+outline our simulations which address the influence of sampling
+uncertainty.
 
@@
 are, to some extent, tracking temperature (see Figure \ref{overview});
-65\% track at least one quantile and 43\% track at least two.
-% Giving an index in units of distance is a good idea; I think it
-% needs some thought, though, because our index is unitless because we
-% divide the units through; I'm not sure the number we get out of this
-% would be directly interpretable like theirs.
-The three temperature indices are, however, the least-tracked of any
-climate variable, while precipitation is the most-tracked (86\% of
-species track at least one of its quantiles). Notably, however, many
-taxa are not tracking climate through time, and many have experienced
-more change than if their distributions had remained static. Just as
-most species are tracking at least one quantile of one climate
-variable, 99.9\% of species have `overshoot' (have a $track_\tau$ less
-than 0; for example becoming colder in a warming region) in at least
-one quantile of at least one climate variable. As we show in Figure
-\ref{overview}, these results are consistent with our null simulations
-where species' range change is a combination of essentially random
-movement and a moderate degree of subsequent climate filtering. Thus
-while our results confirm that, by-and-large, species are tracking
-climate change, the magnitude of this varies across species, taxa, and
-aspects of climate (see Figure \ref{overview}). For example, while
-fungi and insects are, on the whole, broadly tracking the median of
-temperature (median $track_{50\%}$ of 0.46 and 0.64, respectively),
-plants and mammals track temperature (median $track_{50\%}$ of 1.16
-and 1.07, respectively) more poorly than precipitation (median
-$track_{5\%}$ of 0.79 and 0.89, respectively).
+65\% track at least one quantile and 43\% track at least
+two. Surprisingly, our three temperature indices (mean, minimum, and
+maximum temperature) are the least-tracked of any climate variable,
+while precipitation is the most-tracked (86\% of species track at
+least one of its quantiles). Notably, however, [only XXX\% of/no] taxa
+are tracking all quantiles of each climate variable, and the majority
+have `overshot' and experienced more change in one aspect of their
+niche than if their distributions had remained static (XXX\% have at
+least one $track_\tau$ less than 0; for example, becoming colder in a
+warming region). Thus while our results confirm that, by-and-large,
+species are tracking climate change, the degree of tracking varies
+across species, taxa, and aspects of climate (see Figure
+\ref{overview}). For example, while fungi and insects are, on the
+whole, broadly tracking the median of temperature (median
+$track_{50\%}$ of XXX and XXX, respectively), plants and mammals track
+temperature (median $track_{50\%}$ of XXX and XXX, respectively) more
+poorly than precipitation (median $track_{5\%}$ of XXX and XXX,
+respectively).
+% WDP: We need to integrate the idea that species are trading off
+% different aspects of climate. Probably in the PCA section that comes
+% later.
 
@@
 potential drivers of species' ranges and range tracking, we first
-examined the phylogenetic signal of species' climate quantiles. We
-focused on mammals, birds, reptiles, and plants, since we were able to
-find broadly inclusive, dated phylogenies for these taxa. We found
-strong evidence of phylogenetic signal in species' past (median
-Pagel's $\lambda$\supercite{Pagel1999} of 0.60) and current (median
-$\lambda$ of 0.57) distributions, but limited signal in our index
-(median $\lambda$ of $<$0.001). Past climate showed stronger
-phylogenetic signal than our index in 158 of all 180 (4 taxa, 5
+examined the phylogenetic signal of species' present and past climate
+envelopes. We focused on mammals, birds, reptiles, and plants, since
+we were able to find broadly inclusive, dated phylogenies for these
+taxa. We found strong evidence of phylogenetic signal in species' past
+(median Pagel's $\lambda$\supercite{Pagel1999} of XXX) and current
+(median $\lambda$ of XXX) distributions, but limited signal in our
+index (median $\lambda$ of $<$XXX). Past climate showed stronger
+phylogenetic signal than our index in XXX of all 180 (4 taxa, 5
 quantiles, and 9 climate variables) possible taxon-climate-quantile
-combinations (144 for current climate; see Figures \ref{phylo} and
+combinations (XXX for current climate; see Figures \ref{phylo} and
 \ref{traits-signal}). This confirms the known result that species'
 current-day and past ranges are strongly influenced by
-evolution\supercite{Wiens2010}, but suggests that species' realised
-velocity of change (our tracking index) may be associated with
-different factors.
+evolution\supercite{Peterson1999,Wiens2010}, but suggests different
+constraints to contemporary niche tracking.
 
@@
 trait datasets exist for these taxa. Again, we found statistically
-significant associations between traits and past distributions of 191
+significant associations between traits and past distributions of XXX
 of 270 (at $\alpha_{5\%}$; 6 traits, 9 climate variables, and 5
@@
 distributions), but not between the traits and our tracking index (13
-combinations or 5\%; see Figure \ref{traits-signal}). While the traits
-we used are strongly associated with species' climate
-responses\supercite{Wright2004,McCain2014} (hence their strong
-predictive power of species' distributions, but not degree of
-tracking), we emphasise that we could not find sufficient data to
-assess the potential impact of dispersal
-ability\supercite{Schloss2012}. These findings, while surprising, are
-in keeping with emerging evidence that species' traits are a
-relatively poor predictor of change at range edges
-\supercite{Angert2011}.
+combinations or 5\%; see Figure \ref{traits-signal}). The traits we
+used are strongly associated with species' climate
+responses\supercite{Wright2004,McCain2014,Estrada2016} (hence their
+strong predictive power of species' distributions, but not degree of
+tracking), we emphasise that we do not consider species' dispersal
+abilities\supercite{Schloss2012}. This is in part because of limited
+data, but also because of a potential circularity in predicting
+species' observed movements (our tracking index) as a function of
+their measured movement ability (from movement). Our findings, while
+surprising, are in keeping with emerging evidence that species' traits
+are often \supercite{Angert2011,MacLean2017} (but not always
+\supercite{Sunday2015,Pacifici2020}) a poor predictor of change at
+range edges.
 
-\textbf{Species' movement through climate space is disconnected from
-  their position within it.} Given our index is poorly predicted by
-traditionally important variables, we now consider whether index
-values are predictive of each other. It is well known that the axes of
-global climate are not independent and that they are not changing
-independently\supercite{Harris2014,IPCC2014}, and so we would expect
-species' relative tracking of climate to show similar patterns. As we
-show in Figure \ref{pca}, principal components analysis of species'
-current and past climate distributions, which we refer to as their
-climate space, shows strong correlations across climate variables. Yet
-our index does not; species' relative degree of tracking is both of
-much higher dimension (the amount of variance explained by each
-principal component axis is similar) and shows relatively less
-correlation across climates axes . We argue, therefore, that while
-species' positions in climate space show strong associations, species'
-relative tracking of their position in climate space does not. This
-result explains the poor predictive power of species' traits and
-evolutionary history for our index: the ecological and evolutionary
-rules that determine species' climate relationships are
-changing. Thus, in our analysis, we can explain species' current
-distributions only by the degree to which they resemble past
-distributions.
+\textbf{Correlations among climate axes force trade-offs in the
+  magnitude and kind of tracking species can maintain.} It is well
+known that the axes of global climate are not independent and that
+they are not changing independently\supercite{Harris2014,IPCC2014},
+and so we would na\ively expect species' relative tracking of climate
+axes to show similar patterns. As we show in Figure \ref{pca},
+principal components analysis of species' current and past climate
+distributions, which we refer to as their climate space, shows strong
+correlations across climate variables. Yet our index does not:
+species' relative degree of tracking is both of much higher dimension
+(the amount of variance explained by each principal component axis is
+similar) and shows relatively less correlation across climates
+axes. Thus while species' positions in climate space show strong
+associations, species' relative tracking of their position in climate
+space does not. This result suggests that the ecological and
+evolutionary rules that determine species' climate relationships are
+changing, explaining why species' traits and evolutionary history are
+relatively poor predictors of species' niche tracking. That species
+are simultaneously tracking some aspects of climate while overshooting
+others (see above) results from species trading off different,
+correlated aspects of their climatic niche requirements.
 
-There are alternative explanations for our results, of which we
-highlight three: (1) we are ignoring important niche axes, (2) we are
-ignoring biotic interactions, and (3) our index highlights random
-variation in distribution. (1) Our climate variables are not
-all-encompassing, and we ignore other drivers such as geology and
-habitat-type. Yet we do detect strong pattern in current and past
-distributions, suggesting that the absence of pattern in our indices
-is meaningful. (2) Competition and facilitation among species are
-likely strong drivers of species' ability to track
+There are plausible but unlikely alternative explanations for our
+results, of which we highlight three. (1) Our climate variables are
+not all-encompassing, and species may be closely tracking other
+aspects of climate or environment that we do not consider here (\eg,
+geology). Yet we do detect strong pattern in current and past
+distributions, such that the absence of pattern in our tracking
+indices is deserving of explanation and consideration. (2) Competition
+and facilitation among species may mediate species' ability to track
 climate\supercite{Davis1999}. Yet, again, they are surely also likely
-to have driven species' past distributions, and so it is unclear to us
-why only our metric, and not also current and past distributions, is
-so poorly predicted by our trait and phylogenetic data. (3) If species
-were perfectly tracking climate, $current_\tau$ would be equal to
-$past_\tau$ and so our index values could reflect sampling error or
-some other purely stochastic input. Yet such noise, if truly random,
-would mean our empirical index values would be centred at 0 or
-resemble our simulated data, neither or which is the case (see Figure
-\ref{overview}).
+to have driven species' past distributions, begging the question as to
+why species' past and present distributions show such strong trait and
+phylogenetic correlates but their relative tracking do not. (3) If
+species were perfectly tracking climate, $current_\tau$ would be equal
+to $past_\tau$ and so our index values could reflect sampling error or
+some other purely stochastic input. If so, our empirical indices
+(which are bootstrapped to account for sampling differences) should be
+centred at 0, which they are not (see figure \ref{overview}). Further,
+as we show in the supplementary materials, we find strong impacts of
+non-random change even when holding sampling constant, and non-random
+associations and trade-offs in species' niche-tracking.
 
@@
 for example, `no-analogue' conditions generated previously-unseen
-species assemblages \supercite{Williams2007}. Equally, our observation
-of idiosyncrasy in species' degree of tracking is consistent with
-species' population crashes (and so extinctions) becoming less
-predictable during previous mass extinction events
-\supercite{Levinton2001,Jablonski2004}. We call upon others to examine
-the additional drivers of our metric of relative change, and to extend
-its definition to include more nuanced definitions of climate
-space. We predict, however, that if the current mass extinction event
-continues, it is likely that patterns of idiosyncrasy among species'
-declines and distributional changes will become more common.
+species assemblages \supercite{Williams2007,Veloz2012}. Species simply
+cannot track all climate axes of their niches simultaneously, and the
+idiosyncrasies among species' different tracking indices likely
+reflects trade-offs among niche axes.  Species' climate-change
+responses are also inherent context-dependent, and different
+ecosystems and habitats have different drivers
+\supercite{Bertrand2016,Antao2020}, and biotic interactions can both
+mitigate and enhance the impact of climate \supercite{Vandvik2020}.
+Equally, our observation of idiosyncrasy in species' degree of
+tracking is consistent with species' population crashes (and so
+extinctions) becoming less predictable during previous mass extinction
+events \supercite{Levinton2001,Jablonski2004}. Perhaps the key to
+understanding and predicting these seemingly idiosyncratic differences
+in niche tracking will be to understand the distinction between
+species' fundamental and realised niches. By focusing on the
+underlying cut-offs and boundaries that define species' tolerances and
+requirements, we may be able to find pattern among the noise of
+species' responses.
 
@@
 \citeauthor{Harris2014}\supercite{Harris2014}'s nine global
-time-series. We split the data into yearly (averaged) 0.5$^\circ$
-cells, from which we estimated each of our nine climate variables for
-each species observation using
+time-series of mean, minimum, maximum temperature, precipitation,
+`rainy-day' frequency, vapour pressure, potential evapotranspiration,
+`frost day' frequency, and cloud cover. Given recent evidence of the
+divergent roles of mean, minimum, and maximum temperature in driving
+local extinctions \supercite{Roman-Palacios2020}, we present track
+indices for all three aspects of temperature. We split the data into
+yearly (averaged) 0.5$^\circ$ cells, from which we estimated each of
+our nine climate variables for each species observation using
 \emph{raster}\supercite{Hijmans2019}. Plant trait data were taken from

### 2020-12-08 src/analysis-pca.R

@@ -4,11 +4,10 @@ source("src/headers.R")
 # Wrapper functions
 .combined.plot <- function(quantile, dims=1:2, clean=5, abs=FALSE){
 
-    index <- na.omit(.simplify.group("bootstrap.index", quantile, clean=clean,  abs=abs)[,1:9])
-    present <- na.omit(.simplify.group("present", quantile, abs=abs)[,1:9])
-    past <- na.omit(.simplify.group("past", quantile, abs=abs)[,1:9])
+    index <- na.omit(.simplify.group("b.track.index", quantile, clean=clean,  abs=abs)[,1:9])
+    present <- na.omit(.simplify.group("pres.dis.pres.env", quantile, abs=abs)[,1:9])
+    past <- na.omit(.simplify.group("past.dis.past.env", quantile, abs=abs)[,1:9])
     
-    #layout(matrix(1:6, nrow=2, byrow=TRUE), widths=c(5,5,5), heights=c(5,1))
     par(mar=c(3.1,3.1,1.1,1.1), mfrow=c(1,3))
     present <- prcomp(present, scale=TRUE)
     past <- prcomp(past, scale=TRUE)
@@ -22,7 +21,7 @@ source("src/headers.R")
     var <- pca$sdev^2
     p.var <- (var / sum(var)) * 100
     par(mar=c(4.5, 4.5, 1.1, 1.1))
-    plot(density, display="filled.contour", cont=seq(10,90,10), xlab=paste0("PC1 (% variance = ",round(p.var[1]),")"), ylab=paste0("PC2 (% variance = ",round(p.var[2]),")"), cex.lab=1.5, cex.axis=1.25)
+    plot(density, display="filled.contour", cont=seq(10,90,10), xlab=paste0("PC1 (% variance = ",round(p.var[1]),")"), ylab=paste0("PC2 (% variance = ",round(p.var[2]),")"), cex.lab=1.5, cex.axis=1.25, main="")
     points(pca$x, pch=20, cex=.5, col=alpha("grey60",.25))
     arrows(rep(0,ncol(pca$x)),rep(0,ncol(pca$x)), pca$rotation[,dims[1]]*scale, pca$rotation[,dims[2]]*scale, length=.1, lwd=3)
     text.locs <- pca$rotation*scale*1.3
@@ -30,21 +29,24 @@ source("src/headers.R")
 }
 
 # Load data
-plants <- readRDS("clean-data/plants-index.RDS")
[…]

### 2020-12-08 src/analysis-violin.R

@@ -2,39 +2,28 @@
 source("src/headers.R")
 
 # Load empirical data
-plants <- readRDS("clean-data/plants-index.RDS")
-fungi <- readRDS("clean-data/fungi-index.RDS")
-insects <- readRDS("clean-data/insects-index.RDS")
-mammals <- readRDS("clean-data/mammals-index.RDS")
-reptiles <- readRDS("clean-data/reptiles-index.RDS")
-birds <- readRDS("clean-data/birds-index.RDS")
-amphibians <- readRDS("clean-data/amphibians-index.RDS")
-
-# Load simulated data
-load("src/sim-range.RData")
-mid <- t(sapply(results, function(x) x[1,]))
-data <- cbind(data, mid)
-data <- data[!is.na(data$bootstrap.index),]
-sim.data <- data[is.finite(data$index),]
+plants <- Filter(Negate(is.character), readRDS("clean-data/rawbin-clnspc-100-TRUEplants-index.RDS"))
+fungi <- Filter(Negate(is.character), readRDS("clean-data/rawbin-clnspc-100-TRUEfungi-index.RDS"))
+insects <- Filter(Negate(is.character), readRDS("clean-data/rawbin-clnspc-100-TRUEinsects-index.RDS"))
+mammals <- Filter(Negate(is.character), readRDS("clean-data/rawbin-clnspc-100-TRUEmammals-index.RDS"))
+reptiles <- Filter(Negate(is.character), readRDS("clean-data/rawbin-clnspc-100-TRUEreptiles-index.RDS"))
+#birds <- readRDS("clean-data/birds-index.RDS")
+birds <- fungi
+amphibians <- Filter(Negate(is.character), readRDS("clean-data/rawbin-clnspc-100-TRUEamphibians-index.RDS"))
 
 # Do the plots
-.outline.plot <- function(climate, quant, clean=NA, se.thresh=5, abs=FALSE, index="bootstrap.index", sims=FALSE){
-    # Prep simulated data
-    #sim.data <- sim.data[as.character(sim.data$quantile)==quant,]
-    warning("Currently only have simulated data for 0.5 quantile and temperature")
-    sim.data <- sim.data[order(sim.data$alpha),]
-    sim.data$index <- sim.data[,index]
-    sim.data$alpha.cat <- ifelse(sim.data$alpha==0.5, "intermediate", ifelse(sim.data$alpha==0, "static", "tracking"))
[…]

### 2020-12-08 src/calculate-index.R

@@ -46,7 +46,7 @@ calc.metric <- function(pres.dis.pres.env, pres.dis.past.env, past.dis.pres.env,
     }
     
     # Calculate observed quantiles and indices
-    quants <- sapply(list(pres.dis.pres.env, pres.dis.past.env, past.dis.pres.env, past.dis.past.env), quantile, probs=quantile, na.rm=TRUE)
+    quants <- setNames(sapply(list(pres.dis.pres.env, pres.dis.past.env, past.dis.pres.env, past.dis.past.env), quantile, probs=quantile, na.rm=TRUE), c("pres.dis.pres.env", "pres.dis.past.env", "past.dis.pres.env", "past.dis.past.env"))
     track.index <- (quants["pres.dis.pres.env"]-quants["past.dis.past.env"]) / (quants["past.dis.pres.env"]-quants["past.dis.past.env"])
     null.index <- (quants["pres.dis.pres.env"]-quants["pres.dis.past.env"]) / (quants["past.dis.pres.env"]-quants["past.dis.past.env"])
 
@@ -89,15 +89,14 @@ calc.metric.parallel <- function(data, quantiles, rasters, taxon, output.dir, bi
         # Do work per raster
         for(i in seq_along(rasters)){
             t <- map.climate(curr.data, rasters[[i]], 1990:2015, 1955:1980, 35, binary)
-            output[1,i,,] <- t(sapply(quantiles, function(x) calc.metric(t$pres.dis.pres.env, t$pres.dis.past.env, t$past.dis.pres.env, t$past.dis.past.env, x)))            
-            curr.data$year <- sample(curr.data$year)
-            t <- map.climate(curr.data, rasters[[i]], 1990:2015, 1955:1980, 35, binary)
+            output[1,i,,] <- t(sapply(quantiles, function(x) calc.metric(t$pres.dis.pres.env, t$pres.dis.past.env, t$past.dis.pres.env, t$past.dis.past.env, x)))
+            s.data <- curr.data; s.data$year <- sample(s.data$year, size=nrow(s.data))
+            t <- map.climate(s.data, rasters[[i]], 1990:2015, 1955:1980, 35, binary)
             output[2,i,,] <- t(sapply(quantiles, function(x) calc.metric(t$pres.dis.pres.env, t$pres.dis.past.env, t$past.dis.pres.env, t$past.dis.past.env, x)))
         }
         return(output)
     }
-    .int.calc(data$scientificName[1], TRUE)
[…]

### 2020-12-08 src/download-gbif.rb

@@ -2,9 +2,10 @@
 
 # Change the location (and its match in `headers.R`) to download GBIF
 # data elsewhere
-
+require 'fileutils'
 Dir.chdir "~/bear-share/gbif/"
 def download_wrapper(gbif_id, taxon, clean_csv=true, clean_zip=true, clean_gbif=false)
+  unless Dir.exists?(taxon) then FileUtils.mkdir(taxon) end
   Dir.chdir taxon
   `wget http://api.gbif.org/v1/occurrence/download/request/#{gbif_id}.zip`
   `unzip #{gbif_id}.zip #{gbif_id}.csv`
@@ -18,11 +19,11 @@ end
 # If you want to use different GBIF inputs for each taxon, replace the
 # IDs below with something different. You will then need to change
 # `calculate-index.R` since the raw data will have changed
-download_wrapper "0034734-180508205500799" "amphibia"
-download_wrapper "0034724-180508205500799" "asco-basidio"
-download_wrapper "0000172-180412121330197" "aves"
-download_wrapper "0034726-180508205500799" "insecta"
-download_wrapper "0000171-180412121330197" "mammalia"
-download_wrapper "0001849-180131172636756" "plantae"
-download_wrapper "0001851-180131172636756" "plantae"
-download_wrapper "0034728-180508205500799" "reptillia"
+download_wrapper("0034734-180508205500799", "amphibia")
+download_wrapper("0034724-180508205500799", "asco-basidio")
+download_wrapper("0000172-180412121330197", "aves")
+download_wrapper("0034726-180508205500799", "insecta")
+download_wrapper("0000171-180412121330197", "mammalia")
+download_wrapper("0001849-180131172636756", "plantae")
+download_wrapper("0001851-180131172636756", "plantae")
+download_wrapper("0034728-180508205500799", "reptillia")

### 2020-12-08 src/headers.R

@@ -122,29 +122,6 @@ prog.bar <- function(x, y){
     data <- data[year > 1955 & year <= 2015,]
     return(data)
 }
-.calc.metric <- function(present, past, projected, quantile=.5, n.boot=999){
-    # Calculate observed quantiles
-    m.present <- quantile(present, quantile, na.rm=TRUE)
-    m.past <- quantile(past, quantile, na.rm=TRUE)
-    m.projected <- quantile(projected, quantile, na.rm=TRUE)
-    index <- (m.present-m.past) / (m.projected-m.past)
-
-    # Bootstrap past and present sampling
-    b.present <- replicate(n.boot, sample(present, length(past), replace=TRUE), simplify=FALSE)
-    b.present <- sapply(b.present, quantile, probs=quantile, na.rm=TRUE)
-    b.past <- replicate(n.boot, sample(past, length(present), replace=TRUE), simplify=FALSE)
-    b.past <- sapply(b.past, quantile, probs=quantile, na.rm=TRUE)
-    b.projected <- replicate(n.boot, sample(projected, length(present), replace=TRUE), simplify=FALSE)
-    b.projected <- sapply(b.projected, quantile, probs=quantile, na.rm=TRUE)
-    b.mad <- mad((b.present-b.past) / (b.projected-b.past))
-    b.index <- median((b.present-b.past) / (b.projected-b.past))
-    
-    # Format and return
-    return(setNames(
-        c(index, b.index, b.mad, quantile,m.present,m.past,m.projected),
-        c("index","bootstrap.index", "mad.index", "quantile","present","past","projected")
-    ))
-}
 .assign.outcome <- function(present, past, projected){
     if(present==past)
         return("track")
@@ -176,8 +153,8 @@ prog.bar <- function(x, y){
     }
     return("ERROR")
 }
-.simplify <- function(data, index, quantile, clean=NA, abs=FALSE){
-    data <- t(data[,quantile,index,])
+.simplify <- function(data, index, quantile, null, clean=NA, abs=FALSE){
+    data <- t(sapply(data, function(x) x[null,,quantile,index]))
     if(!is.na(clean)){
         keep <- rep(TRUE, nrow(data))
         for(i in seq_len(ncol(data)))
@@ -188,15 +165,15 @@ prog.bar <- function(x, y){
         data <- abs(data)
     return(data)
 }
-.simplify.group <- function(index, quantile, clean=NA, abs=FALSE)
[…]

### 2023-01-27 ms/ms.tex
Pre Montpelier

@@
 they are not changing independently\supercite{Harris2014,IPCC2014},
-and so we would na\ively expect species' relative tracking of climate
+and so we would na\i vely expect species' relative tracking of climate
 axes to show similar patterns. As we show in Figure \ref{pca},

### 2023-01-27 src/analysis-phylo.R

@@ -1,5 +1,7 @@
 # Headers
 source("src/headers.R")
+stem <- "clnbin-clnspc-100-TRUE"
+indices <- c("track.index", "b.track.index", "mad.track.index", "null.index",  "b.null.index", "mad.null.index", "pres.dis.pres.env", "pres.dis.past.env", "past.dis.pres.env", "past.dis.past.env")
 
 # Wrapper functions
 .make.comparative.data <- function(data, tree, merge=TRUE){
@@ -15,36 +17,25 @@ source("src/headers.R")
 }
 .phy.sig <- function(c.data, method="lambda")
     sapply(c.data$data, function(x) phylosig(c.data$phy, x, method))
-.dom.niche <- function(c.data)
-    setNames(
-        apply(as.matrix(c.data$data), 1, function(x) which.min(abs(x))),
-        rownames(c.data$data)
-    )
-.phy.sig.wrap <- function(data, tree, index, clean, method="lambda", merge=TRUE, p.val=FALSE){
-    output <- matrix(NA, ncol=9, nrow=5)
-    rownames(output) <- dimnames(data)[[2]]
-    colnames(output) <- dimnames(data)[[1]]
+
+.phy.sig.wrap <- function(stem, taxon, tree, index, null, clean, method="lambda", merge=TRUE, abs=FALSE){
+    data <- Filter(Negate(is.character), readRDS(paste0("clean-data/",stem,taxon,"-index.RDS")))
+    data <- lapply(data, function(x) x[null,,,index])
+    data <- array(unlist(data), dim=c(dim(data[[1]]), length(data)), dimnames=list(rownames(data[[1]]),colnames(data[[1]]),names(data)))
+    output <- matrix(NA, ncol=ncol(data), nrow=nrow(data))
+    rownames(output) <- dimnames(data)[[1]]
+    colnames(output) <- dimnames(data)[[2]]
+    p.output <- output
     
-    for(i in seq_len(nrow(output))){
-        c.data <- .make.comparative.data(.simplify(data, index, rownames(output)[i]), tree, merge)
-        if(p.val){
-            output[i,] <- as.numeric(.phy.sig(c.data)[2,])
-        } else {
-            output[i,] <- as.numeric(.phy.sig(c.data)[1,])
-        }
-            
+    for(i in seq_len(ncol(output))){
[…]

### 2023-01-27 src/analysis-traits.R

@@ -1,5 +1,6 @@
 # Headers
 source("src/headers.R")
+stem <- "clnbin-clnspc-100-TRUE"
 
 # Wrapper functions
 .univariate <- function(data, explanatory){
@@ -11,9 +12,11 @@ source("src/headers.R")
     colnames(output) <- names(data); rownames(output) <- c("r","p")
     return(output)
 }
-.combine.data <- function(index, quant, clean=10, abs=FALSE){
-    data <- .simplify.group(index, quant, clean, abs)
+.combine.data <- function(stem, index, quant, null, clean=10, abs=FALSE){
+    c(data,metadata) %<-% .load.indices(stem)
+    data <- as.data.frame(.simplify(data, index, quant, null, clean, abs))
     data$species <- rownames(data)
+    data$taxon <- metadata$taxon[match(data$species, metadata$species)]
     
     data$bm.mamm <- neotoma$log.mass[match(data$species, neotoma$binomial)]
     data$bm.mamm[data$taxon != "mammals"] <- NA
@@ -26,9 +29,8 @@ source("src/headers.R")
     data$log.Amass <- glopnet$log.Amass[match(data$species, glopnet$Species)]
     return(data)    
 }
-.corr.mat <- function(index, quant, clean, abs=FALSE, p.val=FALSE){
-    data <- .combine.data(index, quant, clean, abs)
-
+.corr.mat <- function(stem, index, quant, null, clean, abs=FALSE, p.val=FALSE){
+    data <- .combine.data(stem, index, quant, null, clean, abs)
     c.var <- setNames(
         c("clouds","frost","evapo-trans.","precipitation","min(temp)","mean(temp)","max(temp)","vapour","rainy day"),
         c("cld","frs","pet","pre","tmn","tmp","tmx","vap","wet")
@@ -54,13 +56,13 @@ source("src/headers.R")
     }
     return(mat)
 }
-.plot.corr.mat <- function(index, comparison, quant, clean, abs=FALSE){
-    index.mat <- t(.corr.mat(index, quant, clean, abs))
-    index.mat.p <- t(.corr.mat(index, quant, clean, abs, p.val=TRUE))
+.plot.corr.mat <- function(stem, index, comparison, quant, null, clean, abs=FALSE){
[…]

### 2023-01-27 src/analysis-violin.R

@@ -1,30 +1,25 @@
 # Headers
 source("src/headers.R")
-
-# Load empirical data
-plants <- Filter(Negate(is.character), readRDS("clean-data/rawbin-clnspc-100-TRUEplants-index.RDS"))
-fungi <- Filter(Negate(is.character), readRDS("clean-data/rawbin-clnspc-100-TRUEfungi-index.RDS"))
-insects <- Filter(Negate(is.character), readRDS("clean-data/rawbin-clnspc-100-TRUEinsects-index.RDS"))
-mammals <- Filter(Negate(is.character), readRDS("clean-data/rawbin-clnspc-100-TRUEmammals-index.RDS"))
-reptiles <- Filter(Negate(is.character), readRDS("clean-data/rawbin-clnspc-100-TRUEreptiles-index.RDS"))
-#birds <- readRDS("clean-data/birds-index.RDS")
-birds <- fungi
-amphibians <- Filter(Negate(is.character), readRDS("clean-data/rawbin-clnspc-100-TRUEamphibians-index.RDS"))
+stem <- "clnbin-clnspc-100-TRUE"
 
 # Do the plots
-.outline.plot <- function(climate, quant, clean=NA, se.thresh=5, abs=FALSE, index="b.track.index", sims=FALSE){
+.outline.plot <- function(stem, climate, quant, clean=NA, se.thresh=5, abs=FALSE, index="b.track.index", sims=FALSE){
     # Load and match data
-    sim.index <- .simplify.group(index, quantile=quant, null="spp.year.shuffle", clean=clean, abs=abs)
-    index <- .simplify.group(index, quantile=quant, null="observed", clean=clean, abs=abs)
+    c(data,metadata) %<-% .load.indices(stem)
+    sim.index <- as.data.frame(.simplify(data, index, quant, "spp.year.shuffle", clean, abs))
+    sim.index$taxon <- metadata$taxon[match(rownames(sim.index), metadata$species)]
+    index <- as.data.frame(.simplify(data, index, quant, "observed", clean, abs))
+    index$taxon <- metadata$taxon[match(rownames(index), metadata$species)]
     data <- data.frame(index=index[,climate], taxon=index$taxon)
     sim.data <- data.frame(index=sim.index[,climate], taxon=sim.index$taxon)
+    
     # Prepare colours and order for plotting
[…]

### 2023-01-27 src/headers.R

@@ -39,6 +39,8 @@ if(!require("png")){
     install.packages("png"); library(png)}
 if(!require("grid")){
     install.packages("grid"); library(grid)}
+if(!require("zeallot")){
+    install.packages("zeallot"); library(zeallot)}
 
 ####################################
 # Global settings ##################
@@ -165,23 +167,49 @@ prog.bar <- function(x, y){
         data <- abs(data)
     return(data)
 }
-.simplify.group <- function(index, quantile, null="observed", clean=NA, abs=FALSE)
+.simplify.group <- function(stem, index, quantile, null="observed", clean=NA, abs=FALSE)
     return(rbind(
-        data.frame(.simplify(plants, index, quantile, null, clean, abs), taxon="plants"),
-        data.frame(.simplify(fungi, index, quantile, null, clean, abs), taxon="fungi"),
-        data.frame(.simplify(insects, index, quantile, null, clean, abs), taxon="insects"),
-        data.frame(.simplify(mammals, index, quantile, null, clean, abs), taxon="mammals"),
-        data.frame(.simplify(reptiles, index, quantile, null, clean, abs), taxon="reptiles"),
-        data.frame(.simplify(amphibians, index, quantile, null, clean, abs), taxon="amphibians"),
-        data.frame(.simplify(birds, index, quantile, null, clean, abs), taxon="birds")
+        data.frame(.simplify(.load.taxon(stem,"plants"), index, quantile, null, clean, abs), taxon="plants"),
+        data.frame(.simplify(.load.taxon(stem,"fungi"), index, quantile, null, clean, abs), taxon="fungi"),
+        data.frame(.simplify(.load.taxon(stem,"insects"), index, quantile, null, clean, abs), taxon="insects"),
+        data.frame(.simplify(.load.taxon(stem,"mammals"), index, quantile, null, clean, abs), taxon="mammals"),
+        data.frame(.simplify(.load.taxon(stem,"birds"), index, quantile, null, clean, abs), taxon="birds"),
+        data.frame(.simplify(.load.taxon(stem,"reptiles"), index, quantile, null, clean, abs), taxon="reptiles"),
+        data.frame(.simplify(.load.taxon(stem,"amphibians"), index, quantile, null, clean, abs), taxon="amphibians")
     ))
-.simplify.quantile <- function(index, var, clean=NA, abs=FALSE)
[…]

### 2023-01-27 src/calculate-index.R
Missing CRU before

@@ -2,136 +2,137 @@
 source("src/headers.R")
 
 # Link past and present distributions with past and present climate
-map.climate <- function(gbif.data, env, pres.years, past.years, yr.diff, binary){
-    .get <- function(d, e){
-        if(nrow(d)==0)
-            return(numeric(0))
-        if(binary){
-            cells <- mask(e, rasterize(d[,.(decimalLongitude,decimalLatitude)], e, fun="first"))@data@values
-            return(cells[!is.na(cells)])
-        }
-        points <- extract(e, d[,.(decimalLongitude,decimalLatitude)])
-        return(points[!is.na(points)])
+map.climate <- function(gbif.data, env, pres.years, past.years, yr.diff, binary, n.boots){
+  .get <- function(d, e){
+    if(nrow(d)==0)
+      return(numeric(0))
+    if(binary){
+      cells <- mask(e, rasterize(d[,.(decimalLongitude,decimalLatitude)], e, fun="first"))@data@values
+      return(cells[!is.na(cells)])
     }
-    # Pre-allocate
-    pres.dis.pres.env <- pres.dis.past.env <- vector("list", length(pres.years)) 
-    past.dis.pres.env <- past.dis.past.env <- vector("list", length(past.years))
+    points <- extract(e, d[,.(decimalLongitude,decimalLatitude)])
+    return(points[!is.na(points)])
+  }
+  # Pre-allocate
+  pres.dis.pres.env <- pres.dis.past.env <- vector("list", length(pres.years)) 
+  past.dis.pres.env <- past.dis.past.env <- vector("list", length(past.years))
 
-    # Extract
-    for(i in seq_along(pres.years)){
-        pres.dis.pres.env[[i]] <- .get(gbif.data[year==pres.years[i]], env[[pres.years[i]-1900]])
-        pres.dis.past.env[[i]] <- .get(gbif.data[year==pres.years[i]], env[[pres.years[i]-1900-yr.diff]])
-    }
-        
-    for(i in seq_along(past.years)){
-        past.dis.pres.env[[i]] <- .get(gbif.data[year==past.years[i]], env[[past.years[i]-1900+yr.diff]])
-        past.dis.past.env[[i]] <- .get(gbif.data[year==past.years[i]], env[[past.years[i]-1900]])
-    }
-    
-    # Reformat and return
-    return(list(
[…]

### 2023-01-27 src/headers.R

@@ -62,10 +62,11 @@ if(options("mc.cores")==24)
     warning("Assuming you have 24 cores on your computer: if not, change `headers.R`")
 
 data.options <- as.data.frame(expand.grid(
-    min.records=c(500, 1000),
-    clean.binomial=c(TRUE,FALSE),
-    clean.spatial=c(TRUE,FALSE),
-    binary=c(TRUE, FALSE)
+  min.records=c(500, 1000),
+  clean.binomial=c(TRUE,FALSE),
+  clean.spatial=c(TRUE,FALSE),
+  binary=c(TRUE, FALSE),
+  n.boots=99
 ))
 
 data.options$output.dir <- with(data.options,
